### 2024-05-03 Paper_Endogenous.tex
Updates from Overleaf

\documentclass[a4paper]{article}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=3cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
%% Useful packages
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{bbm}
\usepackage{setspace}
\usepackage{footmisc}
\usepackage{pdflscape}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{bm}

\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{url}

\renewcommand\footnotelayout{\fontsize{10}{12}\selectfont}

\interfootnotelinepenalty=10000

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\providecommand{\C}{\mathbb{C}}

\theoremstyle{definition}
\newtheorem{defin}{Definição}

\theoremstyle{plain}
\newtheorem{theorem}[defin]{Teorema}
\newtheorem{corollary}[defin]{Corolário}

\linespread{2}
\title{Rawls For Electricity Planning Models}

\author{Lauren Beatty \thanks{Environmental Defense Fund\\  E-mail: lbeatty1@edf.org \hspace{.5cm}Website: \href{https://lbeatty1.github.io}{https://lbeatty1.github.io}}}

\date{\today}

\begin{document}
\maketitle
\begin{center}
    PRELIMINARY DRAFT - PLEASE DO NOT CITE OR DISTRIBUTE
\end{center}

\begin{abstract}
Renewed interest in the distributional impacts of policies require going beyond typical methods of analysis such as benefit-cost analysis.  In this paper we show how to use least-cost electricity planning models to design equitable systems by turning their logic upside-down.  Rather than using the models to minimize costs, we provide the model with a min-max objective function on distributional health effects of electricity generation and subject the solution to a total cost constraint.  This provides a model result that 
\end{abstract}

\newpage
\section{Introduction}
Climate change mitigation will require deep cuts to greenhouse gas emissions from the electricity sector.  Least-cost electricity planning models are powerful tools that allow policymakers and capacity planners to select portfolios of generation technologies that meet reliability requirements and emissions goals.

There is a large body of work on the health effects of air pollution, and a large body of work on the effects of electricity generation on air pollution. 
Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acture lower respiratory illness, cerebrovascular disease, ischaemic heart disease, and lung cancer. \cite{Lelieveld2015TheScale} estimate that power generation is linked with 31$\%$ of premature mortality caused by outdoor air pollution in the U.S.

There is also work that has shown that air pollution exposures vary drastically across racial and socioeconomic groups.   \cite{Thind2019FineGeography} find that Black and non-Hispanic white people are exposed to more pollution than other groups and that exposures for low-income people are higher than for high-income people, even after controlling for race.  This highlights the importance for developing planning methods that alleviate rather than reinforce disproportionate burdens.

Related work has included co-benefits involved with reducing electricity generation from fossil fuels in the objective function \citet{Sergi2020OptimizingBenefits}.
Related work has also been used to project how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}. However, in this paper, rather than modifying the objective function to include health effects, or use model outputs to project health changes, we endogenize pollution exposure within the electricity model in a way that explicity considers distributional effects.  This allows us to use the model to produce results that don't unfairly saddle certain groups with disproportionate pollution exposure.  It also can be used to answer questions such as "How much can pollution exposure be reduced for $\$100$ billion?"

In future work, we hope to apply a similar methodology to jobs.  This will help policymakers and planners design electricity systems that avoid causing disruptive employment declines in certain regions. Previous work has linked declines in employment from  coal extraction to negative impacts such as decreases in home values, increases in poverty, and increases in credit utilization and delinquency. Moreover, these studies also suggest that even residents who are not employed by the industry also suffer the negatives consequences associated with the industry's decline (\citet{Kraynak2023TheCountry}, \citet{Blonz2023TheFuels}).

\section{Data}
\subsection{Emissions Data}
Emissions data comes from two sources: EPA's Clean Air Markets Program Data (CAMD) and EPA's National Emissions Inventory (NEI).  The CAMD was created to track compliance with clean air programs, and thus, tracks emissions of CO$_2$, NO$_x$, and SO$_2$.  The NEI was created to track criteria pollutants, criteria precursors, and hazardous air pollutants from \textit{all} sources.  I get emissions of NH$_3$, VOC, and PM2.5 from the NEI.

\subsection{Power Sector Modelling Inputs}
Stuff about PowerGenome.

\section{Methods}

The goal of this paper is to use least-cost electricity planning models to design power systems that produce equitable distributions of air pollution exposure.  The modelling can be broken into the following parts: modelling air pollution from electricity generation and modifying Switch to accommodate our min-max objective function.

\subsection{Air Pollution Modelling}
The end product necessary to endogenize air pollution within electricity planning models are a set of coefficients that describe the marginal effect of a unit of electricity generation on population-weighted mean pollution exposure for different racial groups (though you could also apply the same logic and methodology to differing socio-economic groups more broadly).

To do this we employ a reduced-form source receptor matrix that that describe the marginal effect of a unit of pollution on downwind concentration of pollution exposure called the InMap Source Receptor Matrix (ISRM).  InMap is a reduced-form emissions to exposure model that uses annual-average parameters to reduce computation complexity relative to a time-resolved chemical transport model. InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5.

\section{Conclusion}

\begin{singlespace}
\newpage
\bibliographystyle{jpe}
%\bibliographystyle{econometrica}
\bibliography{references.bib}
\end{singlespace}

\end{document}

### 2024-05-03 Slides_5_3_2024.tex
Updates from Overleaf

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{latexsym,amsmath,xcolor,multicol,booktabs,calligra, animate, subfig}
\usepackage{graphicx,pstricks,listings,stackengine, bbm}
\usepackage{tikz}
\usetikzlibrary{fit,tikzmark}   
\usepackage{booktabs,cellspace}
\usepackage{color, colortbl}
\usepackage{hyperref}
\usepackage{color}
%Information to be included in the title page:
\title{Designing Equitable Electricity Systems}
\author{Lauren Beatty\\ \href{mailto:lbeatty@edf.org}{lbeatty@edf.org}\\
\vspace{.5cm}\\
Matthias Fripp\\
\href{mailto:mfripp@edf.org}{mfripp@edf.org}}
\institute{Environmental Defense Fund}
\date{May 6, 2024}

%% Tikz
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{start} = [rectangle, rounded corners, 
minimum width=3cm, 
minimum height=1cm,
text centered, 
draw=black, 
fill=orange!30]

\tikzstyle{process} = [rectangle, 
minimum width=3cm, 
minimum height=1cm, 
text centered, 
text width=3cm, 
draw=black, 
fill=blue!30]

\tikzstyle{stop} = [rectangle, rounded corners, 
minimum width=3cm, 
minimum height=1cm,
text centered, 
draw=black, 
fill=green!30]

\tikzstyle{arrow} = [thick,->,>=stealth]

\logo{\includegraphics[height=1cm]{EDF_color_webRGB_2022 (2).png}\hspace*{.03\paperwidth}\vspace{.01\paperwidth}}

%Color Pallete
\definecolor{EDFblue}{RGB}{0.0, 51, 204}
\definecolor{EDFgreen}{RGB}{0, 153, 51} 
\definecolor{EDFlightgreen}{RGB}{161, 226, 20}
\definecolor{EDFcyan}{RGB}{51,204,255}
\setbeamercolor{structure}{fg=EDFblue} 
\setbeamercolor{button}{bg=EDFblue,fg=white}
\setbeamertemplate{navigation symbols}{}
\newcommand{\nologo}{\setbeamertemplate{logo}{}} % command to set the logo to nothing

\hypersetup{colorlinks,linkcolor=EDFblue,urlcolor=EDFblue}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Introduction}
\textbf{Main Question:  How do we design electrical systems that center equity and justice \textit{without} sacrificing efficiency?}
\vspace{1cm}

\paragraph{\textbf{Previous Paradigm}}
\begin{itemize}
    \item Potentially harmful and polluting infrastructure is sited disproportionately near poor and marginalized communities.
    \item Planning decisions are informed by \textit{least cost} models.
    \item Communities are left in the lurch by technological changes.
\end{itemize}
\end{frame}

\begin{frame}{Background on Least-Cost Electricity Planning Models}
\begin{center}
\begin{tikzpicture}[node distance=2cm]
\node (start) [start, text width=5cm] {\shortstack[l]{• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
• Constraints}};

\node(obj)[process, below of=start, text width=8cm]{\shortstack[l]{$\min \sum \text{Costs}$ \\ \textit{ 
   s.t.   }\text{Everyone gets electricity and nothing breaks }}};

\node (result) [stop, below of=obj] {\shortstack[l]{• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};

\draw [arrow] (start) -- (obj);
\draw [arrow] (obj) -- (result);

\end{tikzpicture}
\end{center}
\end{frame}

{\nologo
\begin{frame}{Previous Attempts to Shoehorn Justice Into these Models}
\begin{center}
\begin{tikzpicture}[node distance=2cm]

\node (start) [start, text width=5cm] {\shortstack[l]{• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
• Constraints}};

\node(obj)[process, below of=start, text width=8cm]{\shortstack[l]{$\min \sum \text{Costs} + \text{Monetized Air Pollution Impacts}$ \\ \textit{ 
   s.t.   }\text{Everyone gets electricity and nothing breaks }}};

\node (result) [stop, below of=obj] {\shortstack[l]{• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};

\draw [arrow] (start) -- (obj);
\draw [arrow] (obj) -- (result);

\end{tikzpicture}
\end{center}

\begin{itemize}
    \item Monetized impacts don't consider unjust distributional consequences.
    \item Equity weights can be difficult to generate and defend.
\end{itemize}

\end{frame}}

{\nologo
\begin{frame}{Our Approach}
    \begin{center}
\begin{tikzpicture}[node distance=2cm]

\node (start) [start, text width=5cm] {\shortstack[l]{• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
• Constraints}};

\node (air) [start, text width=4cm, right of= start, xshift=3cm] {\shortstack[l]{• Air Pollution Model\\
• Emissions Data}};

\node(obj)[process, below of=start, text width=8cm, xshift=0cm]{\shortstack[l]{$C^* = \min \sum \text{Costs}$ \\ \textit{ 
   s.t.   }\text{Everyone gets electricity and nothing breaks }}};

\node(obj2)[process, below of=obj, text width=8cm, xshift=0cm]{\shortstack[l]{$\min \max\{\text{PollutionExp}_1,...,\text{PollutionExp}_N\}$ \\ \textit{ 
   s.t.   }\text{Everyone gets electricity and nothing breaks,} \\
    $\text{      }\sum \text{Costs}<= C^*+Threshold$ }};

\node (result) [stop, below of=obj2] {\shortstack[l]{• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};

\draw [arrow] (start) -- (obj);
\draw [arrow] (air) |- (obj2);
\draw [arrow] (obj) -- (obj2);
\draw [arrow] (obj2) -- (result);

\end{tikzpicture}
\end{center}
\end{frame}}

\begin{frame}{A Point About Calculus}
    \includegraphics[width=0.48\textwidth]{Figures/Slide Pictures/Transmission1.png} \includegraphics[width=0.48\textwidth]{Figures/Slide Pictures/Transmission2.png}

    The effect of deviations from the least-cost optimum will depend on the shape of the objective function.  This is an \textit{empirical} question!\\
    \vspace{.5cm}
    $\Rightarrow$ Perhaps you can get large improvements in health or jobs at only modest increases in cost.
\end{frame}

\begin{frame}{Our Approach}
    \begin{enumerate}
        \item Use EPA emissions data to calculate emissions rates (Pounds per MWh)
        \item Use InMap Source-Receptor Matrix to calculate how one MWh of generation for each resource affects population-weighted average exposure for different groups.
        \item Use Switch to calculate least cost planning costs.
        \item Modify and run switch to minimize air pollution exposure subject to a cost constraint.
    \end{enumerate}

\paragraph{Benefits}
\begin{itemize}
    \item Avoids the use of equity weights while still centering distributional concerns.
    \item Still finds low-cost solutions.
    \item Generalizable framework
\end{itemize}
\end{frame}

\begin{frame}{Thanks!}
\label{Thanks}

    Please feel free to contact me at:
    \href{mailto:lbeatty@edf.org}{lbeatty@edf.org}\\
    \vspace{1cm}

\begin{frame}{Bonus Slides}

\end{frame}

\end{frame}

\end{document}

### 2024-05-03 LiteratureReview.tex

--- 2024-05-03 Paper_Endogenous.tex
@@
 \linespread{2}
-\title{Rawls For Electricity Planning Models}
+\title{Air Quality and Electricity Generation: Literature Review}
 
@@
 
-\begin{abstract}
-Renewed interest in the distributional impacts of policies require going beyond typical methods of analysis such as benefit-cost analysis.
-In this paper we show how to use least-cost electricity planning models to design equitable systems by turning their logic upside-down.
-Rather than using the models to minimize costs, we provide the model with a min-max objective function on distributional health effects of electricity generation and subject the solution to a total cost constraint.
-This provides a model result that
-\end{abstract}
-
 \newpage
 \section{Introduction}
-Climate change mitigation will require deep cuts to greenhouse gas emissions from the electricity sector.
-Least-cost electricity planning models are powerful tools that allow policymakers and capacity planners to select portfolios of generation technologies that meet reliability requirements and emissions goals.
 
-There is a large body of work on the health effects of air pollution, and a large body of work on the effects of electricity generation on air pollution. 
-Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acture lower respiratory illness, cerebrovascular disease, ischaemic heart disease, and lung cancer.
-\cite{Lelieveld2015TheScale} estimate that power generation is linked with 31$\%$ of premature mortality caused by outdoor air pollution in the U.S.
+There is a rich literature on the effects of electricity generation on pollution and the effects of pollution on human health and well-being.
 
-There is also work that has shown that air pollution exposures vary drastically across racial and socioeconomic groups.
-\cite{Thind2019FineGeography} find that Black and non-Hispanic white people are exposed to more pollution than other groups and that exposures for low-income people are higher than for high-income people, even after controlling for race.
-This highlights the importance for developing planning methods that alleviate rather than reinforce disproportionate burdens.
+In addition, there has been an increasing focus on how researchers and public policy makers can consider the holistic, distributional effects of policies.
 
-Related work has included co-benefits involved with reducing electricity generation from fossil fuels in the objective function \citet{Sergi2020OptimizingBenefits}.
-Related work has also been used to project how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}.
-However, in this paper, rather than modifying the objective function to include health effects, or use model outputs to project health changes, we endogenize pollution exposure within the electricity model in a way that explicity considers distributional effects.
-This allows us to use the model to produce results that don't unfairly saddle certain groups with disproportionate pollution exposure.
-It also can be used to answer questions such as "How much can pollution exposure be reduced for $\$100$ billion?"
+Lelieveld et al.  \citep{Lelieveld2015TheScale}
+- air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acture lower respiratory illness (ALRI), cerebrovascular disease (CEV), ischaemic heart disease (IHD), and lung cancer.
+- \textit{We have calculated premature mortality linked to CEV, COPD, IHD and LC for adults 30 years old, and ALRI for infants ,5 years old (Table 1 and Extended Data Tables 1 and 2).
+Our estimate of the globalPM2.5 related mortality in 2010 is 3.15 million people with a 95confidence interval (CI95) of 1.52–4.60 million.
+The main causes are CEV (1.31 million) and IHD (1.08 million), and secondary causes are COPD (374 thousand), ALRI (230 thousand) and LC (161 thousand)}
+- In the U.S., power generation is responsible for 31$\%$ of premature mortality linked to outdoor air pollution
+- Calc deltaMort = y0 [(RR01)/RR]Pop
+- RR=exp(b(X-X0)) where b is concentration response coefficient
 
-In future work, we hope to apply a similar methodology to jobs.
-This will help policymakers and planners design electricity systems that avoid causing disruptive employment declines in certain regions.
-Previous work has linked declines in employment from  coal extraction to negative impacts such as decreases in home values, increases in poverty, and increases in credit utilization and delinquency.
-Moreover, these studies also suggest that even residents who are not employed by the industry also suffer the negatives consequences associated with the industry's decline (\citet{Kraynak2023TheCountry}, \citet{Blonz2023TheFuels}).
+\subsection{InMap}
+\citep{Tessum2017InMAP:Interventions}
+- Annual average changes in primary and secondary PM2.5
+- Eulerian Chemical Transportation Models are computationally expensive and time consuming
+- Uses information from CTM along with simplifying assumptions
+- Variable grid to focus on human exposures
+- Method
 
-\section{Data}
+\subsection{Muller Work - Brief Outline of AP3}
+\citep{}
+
+\section{Description of Air Pollution MIP Modelling}
+
 \subsection{Emissions Data}
@@
 
-\subsection{Power Sector Modelling Inputs}
-Stuff about PowerGenome.
+\subsection{High-level outline}
+\begin{enumerate}
+    \item Join EIA generation data with EPA emissions data to calculate emissions rates in lbs/MWh.
+    \item Calculate emissions from existing plants by taking model-outputted generation and multiplying that by the plant's share of cluster capacity to get a predicted MWh.
+Then multiply that by the plant's emissions rates for all five pollutants.
+    \item New fossil generation outputted by these models is only natural gas.
+For new plants, I start by allocating capacity to both retired and existing sites, with a preference for siting at retired sites.
+\end{enumerate}
 
-\section{Methods}
+https://ehp.niehs.nih.gov/doi/10.1289/EHP9001
 
-The goal of this paper is to use least-cost electricity planning models to design power systems that produce equitable distributions of air pollution exposure.
-The modelling can be broken into the following parts: modelling air pollution from electricity generation and modifying Switch to accommodate our min-max objective function.
-
-\subsection{Air Pollution Modelling}
-The end product necessary to endogenize air pollution within electricity planning models are a set of coefficients that describe the marginal effect of a unit of electricity generation on population-weighted mean pollution exposure for different racial groups (though you could also apply the same logic and methodology to differing socio-economic groups more broadly).
-
-To do this we employ a reduced-form source receptor matrix that that describe the marginal effect of a unit of pollution on downwind concentration of pollution exposure called the InMap Source Receptor Matrix (ISRM).
-InMap is a reduced-form emissions to exposure model that uses annual-average parameters to reduce computation complexity relative to a time-resolved chemical transport model.
-InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5.
-
-\section{Conclusion}
+What are emissions from hydrogen and natgas ccs?
 

### 2024-10-17 Endogenous/Endogenous_makeplots.py
First batch of endogenous results

@@ -9,13 +9,21 @@ import math
 from matplotlib.colors import LinearSegmentedColormap
 
 
+os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
 
-os.chdir('C:/Users/lbeatty/Documents/Lauren_MIP_Contribution/')
+scenario = 'base_short'
+years = [2027, 2030]
+result_dir = f'switch/26-zone/out/foresight/'
+cost_limits = [50,100,200]
 
-scenario = 'current_policies'
-
-coefs = pd.read_csv('MIP_AirPollution/marginal_gen_exposure_coefs_2050.csv', dtype={'Exposure':float}, low_memory=False)
-coefs2 = pd.read_csv('MIP_AirPollution/marginal_gen_exposure_coefs_2050.csv', low_memory=True)
+#read in exposure coefs
+coefs=pd.DataFrame()
+for y in years:
+    coefs = pd.concat([coefs,
+                       pd.read_csv(
+                           f"MIP_Air_OutData/Marginal_Coefficients/{scenario}_marginal_exposure_coefs_{y}.csv",
+                           index_col=0)]
+    )
 
 def update_technology(row):
     if 'naturalgas_ccavg' in row['Cluster']:
@@ -28,20 +36,33 @@ def update_technology(row):
         return 'natural gas ct'
     elif 'coal' in row['Cluster']:
         return 'coal'
+    elif 'biomass' in row['Cluster']:
+        return 'other: biomass, petroleum liquids'
+    elif 'petroleum' in row['Cluster']:
+        return 'other: biomass, petroleum liquids'
     else:
         return np.nan
 
 coefs['technology'] = coefs.apply(update_technology, axis=1)
-coefs2['technology']= coefs2.apply(update_technology, axis=1)
 
-summary=coefs.groupby(['technology', 'Pollutant']).agg({'Exposure':'mean'}).reset_index()
-print(summary.groupby('technology').agg({'Exposure':'sum'}))
+summary=coefs.groupby(['technology', 'Race']).agg({'Exposure':'mean'}).reset_index()
+print(summary)
 
-year='2050'
-emis = gpd.read_file('InMap/MIP_Emissions/marginal_emissions_'+year+'.shp')
+emis = []
+for year in years:
+    shapefile_path = f'MIP_Air_OutData/MIP_Emissions/{scenario}_marginal_emissions_{y}.shp'
+
+    gdf = gpd.read_file(shapefile_path)
+    
[…]

### 2024-10-17 Endogenous/Gen_Emissions_Locations.py

@@ -113,11 +113,12 @@ nei['ann_value']=nei['ann_value']
 nei = nei.pivot(index=['oris_facility_code', 'oris_boiler_id'], columns='poll', values='ann_value').reset_index()
 nei = nei.rename(columns = {'oris_facility_code': 'CAMD_PLANT_ID', 'oris_boiler_id': 'CAMD_UNIT_ID'})
 
-
+print("Finished Reading Data")
 
 #####################
 ## Format columns ###
 #####################
+print("Formatting Data")
 existing_gen_units = existing_gen_units.rename(columns={'generator_id': 'EIA_GENERATOR_ID', 'plant_id_eia':'EIA_PLANT_ID'})
 
 ## EPA-EIA crosswalk
@@ -230,6 +231,7 @@ existing_gen_units['NET_GEN_GENERATOR'] = existing_gen_units['NET_GEN_GENERATOR'
 
 #looks like there's broad agreement between camd and nei
 #fill in missing camd data with nei data
+print("Calculating Emissions Rates")
 
 existing_gen_units['nox_tons'] = existing_gen_units['nox_tons'].combine_first(existing_gen_units['NOX_nei_tons'])
 existing_gen_units['so2_tons'] = existing_gen_units['so2_tons'].combine_first(existing_gen_units['SO2_nei_tons'])
@@ -292,7 +294,7 @@ existing_gen_units['nh3rate_imputed'] = existing_gen_units['nh3_rate'].combine_f
 ############################################################################################
 ### 5. Calculate marginal emissions by multiplying rates by that generator's share of capacity for its cluster.
 ############################################################################################ 
-
+print("Calculating Cluster Emissions Rates")
 column = 'capacity_mw'
 
 technology_year_total = existing_gen_units.groupby(['Resource', 'planning_year']).agg({column : 'sum'}).reset_index()
@@ -414,13 +416,16 @@ print(emissions.shape[0])
 
 
 # ## 7. Split the emissions output by planning year and write to shapefiles.
-for year,input in year_inputs.items():
+print("Writing emissions data to shapefiles")
+for input,year in year_inputs.items():
     emissions_temp = emissions[emissions['planning_year']==year]
     
     emissions_temp.columns = ['Longitude', 'Latitude', 'Resource', 'planning_year', 'NOx', 'SOx', 'PM2_5', 'VOC', 'NH3']
[…]

### 2024-10-17 Endogenous/InMap_SR_GenConstraints.py

@@ -1,5 +1,3 @@
-#%%
-
 from __future__ import (absolute_import, division, print_function, unicode_literals)
 from builtins import * 
 from io import BytesIO, TextIOWrapper
@@ -18,10 +16,10 @@ import geopandas as gpd
 import s3fs
 import os
 
+## Code lifted (almost) directly from https://www.inmap.run/blog/2022/12/15/tutorial/
 
-os.chdir('C:/Users/lbeatty/Documents/Lauren_MIP_Contribution/')
+os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
 
-#%%
 def rect(i, w, s, e, n):
     x = [w[i], e[i], e[i], w[i], w[i]]
     y = [s[i], s[i], n[i], n[i], s[i]]
@@ -47,10 +45,11 @@ sr = zarr.open(s3fs.S3Map(url, s3=fs, check=False), mode="r")
 
 p = poly(sr)
 
-#%%
-years = ['2030', '2040', '2050']
+years = ['2045', '2050']
+scenario = 'base_short'
 for year in years:
-    emis = gpd.read_file('InMap/MIP_Emissions/marginal_emissions_'+year+'.shp')
+    filename='MIP_Air_OutData/MIP_Emissions/'+ scenario+ '_marginal_emissions_'+str(year)+'.shp'
+    emis = gpd.read_file(filename)
     fact = 28766.639
 
     df = pd.DataFrame({'Location': range(52411)})
@@ -120,6 +119,6 @@ for year in years:
     exposure_data_collapsed = exposure_data.groupby(['Race', 'Cluster','Pollutant']).agg({'Exposure':'sum'}).reset_index()
     exposure_data_collapsed['year']=year
 
-    exposure_data_collapsed.to_csv('MIP_AirPollution/marginal_gen_exposure_coefs_'+year+'.csv')
+    exposure_data_collapsed.to_csv('MIP_Air_OutData/Marginal_Coefficients/' + scenario+'_marginal_exposure_coefs_'+year+'.csv')
 
 

### 2024-10-17 Endogenous/Endogenous_makeplots.py
newfig

@@ -7,6 +7,8 @@ import numpy as np
 import pandas as pd
 import math
 from matplotlib.colors import LinearSegmentedColormap
+from matplotlib.colors import ListedColormap
+
 
 
 os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
@@ -322,3 +324,26 @@ for y in years:
     ax.legend()
 
     plt.savefig(f'MIP_AirPollution/Figures/EndogenousResults/{scenario}/exposure_by_scenario_{y}.png', dpi=300, bbox_inches='tight')  # DPI for high resolution
+
+##
+## PPF
+
+group_exposure_max = group_exposure.groupby(['SetProduct_OrderedSet_2', 'scenario']).agg({'GroupExposure':'max'}).reset_index()
+group_exposure_max['cost'] = group_exposure_max['scenario'].str.replace(f'{scenario}', '', regex=False)
+group_exposure_max['cost'] = group_exposure_max['cost'].str.replace('_', '', regex=False)
+group_exposure_max['cost'] = pd.to_numeric(group_exposure_max['cost'])
+group_exposure_max.fillna(0, inplace=True)
+
+colors = {2027: 'blue', 2030: 'green'}  # Assign specific colors to each year
+
+for year in group_exposure_max['SetProduct_OrderedSet_2'].unique():
+    subset = group_exposure_max[group_exposure_max['SetProduct_OrderedSet_2'] == year]
+    plt.scatter(subset['cost'], subset['GroupExposure'], color=colors[year], label=str(year))
+
+plt.xlabel('Cost')
+plt.ylabel('Result')
+plt.title('Cost vs Result (Colored by Year)')
+
+plt.legend(title='Year')
+
+plt.savefig(f'MIP_AirPollution/Figures/EndogenousResults/{scenario}/exposure_cost_PPF.png', dpi=300, bbox_inches='tight')  # DPI for high resolution

### 2024-10-18 Slides_5_3_2024.tex

@@
 \begin{tikzpicture}[node distance=2cm]
-\node (start) [start, text width=5cm] {\shortstack[l]{• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
+\node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
 • Constraints}};
@@
 
-\node (result) [stop, below of=obj] {\shortstack[l]{• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};
+\node (result) [stop, below of=obj] {\shortstack[l]{Outputs\\• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};
 
@@
 
-\node (start) [start, text width=5cm] {\shortstack[l]{• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
+\node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
 • Constraints}};
@@
    s.t.   }\text{Everyone gets electricity and nothing breaks,} \\
-    $\text{      }\sum \text{Costs}<= C^*+Threshold$ }};
+    $\text{      }\sum \text{Costs} \leq C^*+Threshold$ }};
 
-\node (result) [stop, below of=obj2] {\shortstack[l]{• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};
+\node (result) [stop, below of=obj2] {\shortstack[l]{Outputs\\• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};
 
@@
 
-\begin{frame}{A Point About Calculus}
-    \includegraphics[width=0.48\textwidth]{Figures/Slide Pictures/Transmission1.png} \includegraphics[width=0.48\textwidth]{Figures/Slide Pictures/Transmission2.png}
-
-    The effect of deviations from the least-cost optimum will depend on the shape of the objective function.  This is an \textit{empirical} question!\\
-    \vspace{.5cm}
-    $\Rightarrow$ Perhaps you can get large improvements in health or jobs at only modest increases in cost.
-\end{frame}
-
 \begin{frame}{Our Approach}
@@
         \item Use Switch to calculate least cost planning costs.
-        \item Modify and run switch to minimize air pollution exposure subject to a cost constraint.
+        \item Modify and run Switch to minimize air pollution exposure subject to a cost constraint.
     \end{enumerate}
@@
 
+\begin{frame}{A Point About Calculus}
+    \includegraphics[width=0.48\textwidth]{Figures/Slide Pictures/Transmission1.png} \includegraphics[width=0.48\textwidth]{Figures/Slide Pictures/Transmission2.png}
+
+    The effect of deviations from the least-cost optimum will depend on the shape of the objective function.  This is an \textit{empirical} question!\\
+    \vspace{.5cm}
+    $\Rightarrow$ Perhaps you can get large improvements in health or jobs at only modest increases in cost.
+\end{frame}
+
 \begin{frame}{Thanks!}
@@
     \vspace{1cm}
+\end{frame}
 
-\begin{frame}{Bonus Slides}
+\begin{frame}{Bonus Slide!}
+    This project started as a side-project to the Model Intercomparison Project -- a project aimed at harmonizing the inputs and comparing multiple open-source electricity planning models.
 
-\end{frame}
+    \vspace{.4cm}
+
+    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2030.jpg}
+    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2040.jpg}
+    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2050.jpg}
+
+\begin{itemize}
+    \item Death rates vary widely across racial groups.
+    \item Aggressive carbon policies reduce pollution exposure the fastest.
+    \item Optimistic -- even under current policies future generations should have much better air quality.
+\end{itemize}
 

### 2024-10-18 Endogenous/Endogenous_makeplots.py
new bar chart

@@ -276,22 +276,49 @@ for tech in ['Wind', 'Solar', 'Naturalgas', 'Coal', 'Electricity']:
             plotdata = plotdata[plotdata['period']==y]
 
             plotdata = plotdata.merge(base, on='model_regi', how='left')
-            plotdata['GenDif'] = plotdata[f'Energy_GWh_{tech}']/plotdata['Base_Energy_GWh']
+            plotdata['GenDif'] = plotdata[f'Energy_GWh_{tech}']-plotdata['Base_Energy_GWh']
+
+            vmin = plotdata['GenDif'].min()
+            vmax = plotdata['GenDif'].max()
 
             fig, ax = plt.subplots(1, 1, figsize=(10, 6))
             plotdata.plot(column='GenDif', 
-                            ax=ax, 
-                            legend=True, 
-                            cmap='OrRd',
-                            edgecolor='black')
+                        ax=ax, 
+                        legend=True, 
+                        cmap='OrRd',
+                        edgecolor='black',
+                        vmin=vmin,  # Set minimum value for the colormap
+                        vmax=vmax,
+                        legend_kwds={'label': 'Generation Difference (GWh)',  # Title for the colorbar
+                                     'orientation': 'vertical'})  # Set maximum value for the colormap
 
-            # Optional: Customize the plot further
             ax.set_title(f'Generation Relative to Base: {tech}', fontsize=15)
-            ax.set_axis_off()  # Turn off the axis for a cleaner look
-
-            # Show the plot
+            ax.set_axis_off() 
             plt.savefig(f'MIP_AirPollution/Figures/EndogenousResults/{scenario}/Dispatch_Relative_{tech}_{y}_{c}.png', dpi=300, bbox_inches='tight')  # DPI for high resolution
 
+
+dispatch = dispatch.groupby(['gen_tech', 'period', 'scenario']).agg({'Energy_GWh_typical_yr':'sum'}).reset_index()
+
+other_techs = dispatch[dispatch['scenario']==scenario]
+other_techs = other_techs[other_techs['Energy_GWh_typical_yr']<1000]
+other_techs = other_techs['gen_tech'].tolist()
+dispatch.loc[dispatch['gen_tech'].isin(other_techs), 'gen_tech'] = 'other'
+
+for year in years:
+    data_year = dispatch[dispatch['period'] == year]
+    
[…]

### 2024-10-18 Paper_Endogenous.tex

@@
 Renewed interest in the distributional impacts of policies require going beyond typical methods of analysis such as benefit-cost analysis.
-In this paper we show how to use least-cost electricity planning models to design equitable systems by turning their logic upside-down.
-Rather than using the models to minimize costs, we provide the model with a min-max objective function on distributional health effects of electricity generation and subject the solution to a total cost constraint.
-This provides a model result that
+In this paper we show how to use least-cost electricity planning models to design equitable systems by modifying their objective functions.
+Rather than using the models to minimize costs, we provide the model with a min-max objective function on the distributional health effects of electricity generation and subject the solution to a total cost constraint.
+This method provides a menu of plans that equitably reduce pollution exposure.
 \end{abstract}
@@
 
+However, these models have historically ignored some of the external costs of electricity generation. 
+ One major external costs is the generation of local air pollutants.
 There is a large body of work on the health effects of air pollution, and a large body of work on the effects of electricity generation on air pollution.
-Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acture lower respiratory illness, cerebrovascular disease, ischaemic heart disease, and lung cancer.
+Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acute lower respiratory illness, cerebrovascular disease, ischaemic heart disease, and lung cancer.
 \cite{Lelieveld2015TheScale} estimate that power generation is linked with 31$\%$ of premature mortality caused by outdoor air pollution in the U.S.
@@
 Related work has included co-benefits involved with reducing electricity generation from fossil fuels in the objective function \citet{Sergi2020OptimizingBenefits}.
-Related work has also been used to project how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}.
-However, in this paper, rather than modifying the objective function to include health effects, or use model outputs to project health changes, we endogenize pollution exposure within the electricity model in a way that explicity considers distributional effects.
-This allows us to use the model to produce results that don't unfairly saddle certain groups with disproportionate pollution exposure.
-It also can be used to answer questions such as "How much can pollution exposure be reduced for $\$100$ billion?"
+or projected how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}.
+However, in this paper, rather than modifying the objective function to include health effects, or use model outputs to project health changes, we endogenize pollution exposure within the electricity model in a way that explicitly considers distributional effects.
+This allows us to use the model to produce results that don't unfairly burden certain groups with disproportionate pollution exposure.
+It also can be used to answer questions such as "How much can pollution exposure be equitably reduced for $\$100$ billion?"
 
-In future work, we hope to apply a similar methodology to jobs.
-This will help policymakers and planners design electricity systems that avoid causing disruptive employment declines in certain regions.
+This method could also be used to help design electricity transition plans that are equitable in other dimensions.
+For example, one concern with the energy transition is that some regions will lose jobs.
+This method could help policymakers and planners design electricity systems that avoid causing disruptive employment declines in certain regions.
 Previous work has linked declines in employment from  coal extraction to negative impacts such as decreases in home values, increases in poverty, and increases in credit utilization and delinquency.
@@
 
-The goal of this paper is to use least-cost electricity planning models to design power systems that produce equitable distributions of air pollution exposure.
-The modelling can be broken into the following parts: modelling air pollution from electricity generation and modifying Switch to accommodate our min-max objective function.
+Least-cost electricity planning models are the gold-standard models used to model capacity expansion in policy.
+However, these models tend to have a narrow focus on minimizing costs -- potentially at the expense of distributional concerns.
+The energy transition will have wide-ranging effects --such as shifts in the spatial distribution of jobs, or changes in air pollution, that are of interest to policy-makers, but are currently unincorporated into least-cost models.
+
+Previous attempts to consider the health impacts of air pollution have monetized air pollution harms, and inserted them into the cost minimization.
+While this is certainly an improvement over not considering these costs at all, this method ignores the fact that inequity in the past has driven current differences in air pollution exposure, and that certain groups may experience higher burdens from air pollution than others.
+
+In this paper, we outline a method to use least-cost electricity planning models to design power systems that produce equitable distributions of air pollution exposure.
+We pull inspiration from Rawls and formulate our model as a min-max problem where the social planner minimizes the maximum pollution exposure subject to a cost constraint, rather than minimizing costs.
+Running multiple iterations of the model at different cost constraints allows us to understand, how much can we equitably improve air quality and at what costs.
 

### 2024-10-18 Paper_Endogenous.tex
Updates from Overleaf

@@
 \usepackage{bm}
+\usepackage{tikz}
+\usetikzlibrary{fit,tikzmark}  
+\usetikzlibrary{shapes.geometric, arrows}
+\usepackage{color, colortbl}
 
@@
 \renewcommand\footnotelayout{\fontsize{10}{12}\selectfont}
+
+\tikzstyle{start} = [rectangle, rounded corners, 
+minimum width=3cm, 
+minimum height=1cm,
+text centered, 
+draw=black, 
+fill=orange!30]
+
+\tikzstyle{process} = [rectangle, 
+minimum width=3cm, 
+minimum height=1cm, 
+text centered, 
+text width=3cm, 
+draw=black, 
+fill=blue!30]
+
+\tikzstyle{stop} = [rectangle, rounded corners, 
+minimum width=3cm, 
+minimum height=1cm,
+text centered, 
+draw=black, 
+fill=green!30]
+
+\tikzstyle{arrow} = [thick,->,>=stealth]
 
@@
 
-\author{Lauren Beatty \thanks{Environmental Defense Fund\\  E-mail: lbeatty1@edf.org \hspace{.5cm}Website: \href{https://lbeatty1.github.io}{https://lbeatty1.github.io}}}
+\author{Lauren Beatty\thanks{Environmental Defense Fund  \hspace{.5cm} E-mail: lbeatty1@edf.org \hspace{.5cm}Website: \href{https://lbeatty1.github.io}{https://lbeatty1.github.io}}\\
+Matthias Fripp\thanks{Environmental Defense Fund}}
 
@@
 Climate change mitigation will require deep cuts to greenhouse gas emissions from the electricity sector.
-Least-cost electricity planning models are powerful tools that allow policymakers and capacity planners to select portfolios of generation technologies that meet reliability requirements and emissions goals.
+Least-cost electricity planning models are powerful tools that help policymakers and capacity planners find portfolios of generation technologies that meet reliability requirements and emissions goals.
 
@@
 
-Related work has included co-benefits involved with reducing electricity generation from fossil fuels in the objective function \citet{Sergi2020OptimizingBenefits}.
+Previous work has included co-benefits involved with reducing electricity generation from fossil fuels in the objective function \citet{Sergi2020OptimizingBenefits}.
 or projected how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}.
@@
 Moreover, these studies also suggest that even residents who are not employed by the industry also suffer the negatives consequences associated with the industry's decline (\citet{Kraynak2023TheCountry}, \citet{Blonz2023TheFuels}).
+
+In this paper, we use
 
@@
 Running multiple iterations of the model at different cost constraints allows us to understand, how much can we equitably improve air quality and at what costs.
+\begin{figure}
+    \begin{center}
+        \begin{tikzpicture}[node distance=2cm]
+        \node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
+        • Constraints}};
+
+        \node(obj)[process, below of=start, text width=8cm]{\shortstack[l]{$\min \sum \text{Costs}$ \\ \textit{ 
+           s.t.   }\text{Everyone gets electricity and nothing breaks }}};
+
+        \node (result) [stop, below of=obj] {\shortstack[l]{Outputs\\• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};
+
+        \draw [arrow] (start) -- (obj);
+        \draw [arrow] (obj) -- (result);
+
+        \end{tikzpicture}
+    \end{center}
+    \caption{Typical Flow of Least-Cost Planning Models}
+    \label{fig:enter-label}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}[node distance=2cm]
+
+\node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
+• Constraints}};
+
+\node (air) [start, text width=4cm, right of= start, xshift=3cm] {\shortstack[l]{• Air Pollution Model\\
+• Emissions Data}};
+
+\node(obj)[process, below of=start, text width=8cm, xshift=0cm]{\shortstack[l]{$C^* = \min \sum \text{Costs}$ \\ \textit{ 
+   s.t.   }\text{Everyone gets electricity and nothing breaks }}};
+
+\node(obj2)[process, below of=obj, text width=8cm, xshift=0cm]{\shortstack[l]{$\min \max\{\text{PollutionExp}_1,...,\text{PollutionExp}_N\}$ \\ \textit{ 
+   s.t.   }\text{Everyone gets electricity and nothing breaks,} \\
+    $\text{      }\sum \text{Costs} \leq C^*+Threshold$ }};
+
+\node (result) [stop, below of=obj2] {\shortstack[l]{Outputs\\• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};
+
+\draw [arrow] (start) -- (obj);
+\draw [arrow] (air) |- (obj2);
+\draw [arrow] (obj) -- (obj2);
+\draw [arrow] (obj2) -- (result);
+
+\end{tikzpicture}
+\caption{Our Modification to the Least-Cost Planning Model}
+\end{center}
+\end{figure}
 
@@
 InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5.
+
+\section{Results}
+
+\begin{figure}
+\centering
+\subcaptionbox{Mean empirical emissions rates by technology in tons per MWh.}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/base_short_emissions_by_technology.png}}%
+\hfill
+\subcaptionbox{Mean coefficients describing how a marginal MWh of generation affects mean group-level PM2.5 exposure.}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/base_short_exposure_by_technology.png}}%
+\hfill
+\caption{Unsurprisingly, coal is far dirtier than other technologies.  However, the discrepancy between technologies is more narrow for exposure.}
+\label{EmissionsExposureCoefs}
+\end{figure}
+In figure \ref{EmissionsExposureCoefs} we plot both average emissions rates, and average exposure coefficients for various technologies.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{}
+    \caption{Caption}
+    \label{fig:enter-label}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7\linewidth]{Figures/EndogenousResults/base_short/exposure_cost_PPF.png}
+    \caption{Here we show the relationship between the objective function (min-max of exposure) and the cost limit constraint.}
+    \label{PPF}
+\end{figure}
 

### 2024-10-18 exposure/setup_exposure_scenarios.py

@@ -1,29 +1,0 @@
-import pandas as pd
-
-dfs = []
-for y in [2030, 2040, 2050]:
-    dfs.append(
-        pd.read_csv(
-            f"https://raw.githubusercontent.com/lbeatty1/MIP_AirPollution/main/marginal_gen_exposure_coefs_{y}.csv",
-            index_col=0,
-        )
-    )
-exposure = (
-    pd.concat(dfs)
-    .groupby(["Cluster", "year", "Race"], as_index=False)["Exposure"]
-    .sum()
-).query("Exposure != 0")
-exposure.columns = [
-    "GENERATION_PROJECT",
-    "PERIOD",
-    "EXPOSURE_GROUP",
-    "group_exposure_coefficient",
-]
-for case in ["base", "current_policies"]:
-    exposure.to_csv(
-        f"in/foresight/{case}/group_exposure_coefficients.csv", na_rep=".", index=False
-    )
-    df = pd.read_csv(f"in/foresight/{case}/financials.csv", na_values=["."])
-    df["system_cost_limit"] = 1000000000000000
-    df.to_csv(f"in/foresight/{case}/financials.{case}_100.csv", na_rep=["."])
-    df.to_csv(f"in/foresight/{case}/financials.{case}_200.csv", na_rep=["."])

### 2024-12-02 Paper_Endogenous.tex

@@
 or projected how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}.
+
+Least-cost electricity planning models are the gold-standard models used to model capacity expansion in policy.
+However, these models tend to have a narrow focus on minimizing costs -- potentially at the expense of distributional concerns.
+The energy transition will have wide-ranging effects --such as shifts in the spatial distribution of jobs, or changes in air pollution, that are of interest to policy-makers, but are currently unincorporated into least-cost models.
+
+Previous attempts to consider the health impacts of air pollution have monetized air pollution harms, and inserted them into the cost minimization.
+While this is certainly an improvement over not considering these costs at all, this method ignores the fact that inequity in the past has driven current differences in air pollution exposure, and that certain groups may experience higher burdens from air pollution than others.
+
 However, in this paper, rather than modifying the objective function to include health effects, or use model outputs to project health changes, we endogenize pollution exposure within the electricity model in a way that explicitly considers distributional effects.
+We pull inspiration from Rawls and formulate our model as a min-max problem where the social planner minimizes the maximum pollution exposure subject to a cost constraint, rather than minimizing costs.
+Running multiple iterations of the model at different cost constraints allows us to understand how much can we equitably improve air quality and at what costs.
+
 This allows us to use the model to produce results that don't unfairly burden certain groups with disproportionate pollution exposure.
@@
 
-In this paper, we use
+This paper makes an important contribution to the environmental justice literature by ...
 
@@
 \subsection{Power Sector Modelling Inputs}
-Stuff about PowerGenome.
+Stuff about PowerGenome?
 
 \section{Methods}
+\subsection{Overview}
 
-Least-cost electricity planning models are the gold-standard models used to model capacity expansion in policy.
-However, these models tend to have a narrow focus on minimizing costs -- potentially at the expense of distributional concerns.
-The energy transition will have wide-ranging effects --such as shifts in the spatial distribution of jobs, or changes in air pollution, that are of interest to policy-makers, but are currently unincorporated into least-cost models.
-
-Previous attempts to consider the health impacts of air pollution have monetized air pollution harms, and inserted them into the cost minimization.
-While this is certainly an improvement over not considering these costs at all, this method ignores the fact that inequity in the past has driven current differences in air pollution exposure, and that certain groups may experience higher burdens from air pollution than others.
-
-In this paper, we outline a method to use least-cost electricity planning models to design power systems that produce equitable distributions of air pollution exposure.
-We pull inspiration from Rawls and formulate our model as a min-max problem where the social planner minimizes the maximum pollution exposure subject to a cost constraint, rather than minimizing costs.
-Running multiple iterations of the model at different cost constraints allows us to understand, how much can we equitably improve air quality and at what costs.
 \begin{figure}
@@
 \subsection{Air Pollution Modelling}
-The end product necessary to endogenize air pollution within electricity planning models are a set of coefficients that describe the marginal effect of a unit of electricity generation on population-weighted mean pollution exposure for different racial groups (though you could also apply the same logic and methodology to differing socio-economic groups more broadly).
+The end product necessary to endogenize air pollution within electricity planning models are a set of coefficients that describe the marginal effect of a unit of electricity generation on population-weighted mean pollution exposure for different racial groups (though you could also apply the same logic and methodology to differing socio-economic groups more broadly, or more fine slices of the population to more closely approximate a true min-max).
 
-To do this we employ a reduced-form source receptor matrix that that describe the marginal effect of a unit of pollution on downwind concentration of pollution exposure called the InMap Source Receptor Matrix (ISRM).
-InMap is a reduced-form emissions to exposure model that uses annual-average parameters to reduce computation complexity relative to a time-resolved chemical transport model.
-InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5.
+Eulerian chemical transportation models are computationally expensive and time consuming, so it's infeasible to embed this type of model within SWITCH.
+Rather we use the InMAP Source-Receptor Matrix (ISRM) which is a reduced-form source-receptor matrix that describes the marginal effect of a unit of pollution on downwind concentrations of pollution exposure.
+
+InMap itself is a reduced-form emissions-to-exposure model that uses annual-average parameters to reduce computation complexity relative to a time-resolved chemical transport model.
+InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs the annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5 \citep{Tessum2017InMAP:Interventions}.
+
+The ISRM was developed to simplify InMAP even further \citep{Goodkind2019Fine-scaleEmissions}.
+Initially, the ISRM was developed to describe the marginal damages from individual sources.
+We simplify the ISRM even further by collapsing it so that, for each source, we save the marginal effect of one MWh of generation on \textit{population-weighted average exposure} by race.
+
+\subsection{Switch}
 
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{}
-    \caption{Caption}
-    \label{fig:enter-label}
+    \subcaptionbox{2027}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/Dispatch_Relative_by_scenario_2027.png}}
+    \subcaptionbox{2030}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/Dispatch_Relative_by_scenario_2030.png}}
+    \caption{Generation by Technology by Scenario}
 \end{figure}
@@
 \section{Conclusion}
+In this paper, we demonstrate how to endogenize air pollution exposure into a least-cost electricity planning model to create equitable reductions in air pollution burdens.
+
+While this method provides a map of how to produce equitable reductions in air pollution, it does not solve for what policies will get
 

### 2024-12-03 Downscaled/Calculate_Model_Emissions.py
lots of new figs

@@ -4,6 +4,7 @@ globals().clear()
 import pandas as pd
 import matplotlib.pyplot as plt
 import os
+import sys
 import numpy as np
 import geopandas as gpd
 import fiona
@@ -12,10 +13,11 @@ import zipfile
 from shapely.geometry import Point
 
 os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
+sys.path.append(os.path.abspath('C:/Users/laure/Documents/Switch-USA-PG/'))
 
 #pull in settings from settings file
 from MIP_AirPollution.Downscaled.settings import *
-model='GenX'
+models=['GenX','TEMOA', 'SWITCH', 'USENSYS']
 
 with open('MIP_AirPollution/Downscaled/emis_detail.csv', 'w') as file:
     pass
@@ -23,579 +25,582 @@ with open('MIP_AirPollution/Downscaled/emis_detail.csv', 'w') as file:
 with open('MIP_AirPollution/Downscaled/emis.csv', 'w') as file:
     pass
 
+for model in models:
+    for key, value in scenario_dictionary.items():
+        scenario = f"{key}"
+        scenario_name_alt = f"{value}"
+        print('running scenario: '+ scenario)
+            
+        ###################################
+        ## Read in EIA and EPA Data #######
+        ###################################
 
-for key, value in scenario_dictionary.items():
-    scenario = f"{key}"
-    scenario_name_alt = f"{value}"
-    print('running scenario: '+ scenario)
+        #EIA-EPA Crosswalk
+        crosswalk = pd.read_csv(eia_datapath +"/epa_eia_crosswalk.csv")
+
+        # EIA 860 for Generator Info
+        eia860 = pd.read_excel(eia_datapath +"/eia860/3_1_Generator_Y2020.xlsx", skiprows=1)
+
+        # EIA923 for Generation Info
+        eia923_fuels = pd.read_excel(eia_datapath +"/eia923/EIA923_Schedules_2_3_4_5_M_12_2020_Final_Revision.xlsx", sheet_name='Page 1 Generation and Fuel Data', skiprows=5)
+        eia923_generators = pd.read_excel(eia_datapath +"/eia923/EIA923_Schedules_2_3_4_5_M_12_2020_Final_Revision.xlsx", sheet_name='Page 4 Generator Data', skiprows=5)
+
+        ### Format Rows and Clean ####
+
+        ## EPA-EIA crosswalk
[…]

### 2024-12-03 Downscaled/InMAP_SR.py

@@ -1,6 +1,3 @@
-globals().clear()
-
-
 from __future__ import (absolute_import, division, print_function, unicode_literals)
 from builtins import * 
 from io import BytesIO, TextIOWrapper
@@ -18,13 +15,15 @@ import pandas as pd
 import geopandas as gpd
 import s3fs
 import os
+import sys
 
 
 os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
+sys.path.append(os.path.abspath('C:/Users/laure/Documents/Switch-USA-PG/'))
 
 #pull in settings from settings file
 from MIP_AirPollution.Downscaled.settings import *
-model='GenX'
+models=['USENSYS']
 
 def rect(i, w, s, e, n):
     x = [w[i], e[i], e[i], w[i], w[i]]
@@ -152,13 +151,20 @@ def run_sr(emis, model, emis_units="tons/year"):
     print("Finished (%.0f seconds)               "%(time.time()-start))
     return ret
 
+for model in models:
+    for scenario in scenario_dictionary:
+        for file, year in year_inputs.items():
+            if not os.path.exists('InMap/MIP_Emissions/'+scenario+'/'+model+'/emissions_'+f"{year}"+'.shp'):
+                print(f"'{scenario}' emissions output doesn't exist... skipping...")
+                continue
 
-for scenario in scenario_dictionary:
-    for file, year in year_inputs.items():
-        emis = gpd.read_file('InMap/MIP_Emissions/'+scenario+'/'+model+'/emissions_'+f"{year}"+'.shp')
-        print(emis.sum())
-        resultsISRM = run_sr(emis, model="isrm")
-        resultsISRM = resultsISRM.set_crs('PROJCS["Lambert_Conformal_Conic",GEOGCS["GCS_unnamed ellipse",DATUM["D_unknown",SPHEROID["Unknown",6370997,0]],PRIMEM["Greenwich",0],UNIT["Degree",0.0174532925199433]],PROJECTION["Lambert_Conformal_Conic_2SP"],PARAMETER["latitude_of_origin",40],PARAMETER["central_meridian",-97],PARAMETER["standard_parallel_1",33],PARAMETER["standard_parallel_2",45],PARAMETER["false_easting",0],PARAMETER["false_northing",0],UNIT["metre",1,AUTHORITY["EPSG","9001"]],AXIS["Easting",EAST],AXIS["Northing",NORTH]]')
[…]

### 2024-12-03 Downscaled/Plot_Output_ISRM.py

@@ -9,34 +9,41 @@ import pandas as pd
 import math
 from matplotlib.colors import LinearSegmentedColormap
 import seaborn as sns
+import sys
 
 
 os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
+sys.path.append(os.path.abspath('C:/Users/laure/Documents/Switch-USA-PG/'))
 
 from MIP_AirPollution.Downscaled.settings import *
-model='GenX'
+models = ['GenX', 'SWITCH', 'TEMOA', 'USENSYS']
 
 
 #%%
 ## Read in all data
 emissions = pd.DataFrame()
 
-for scenario in scenario_dictionary:
-    print('reading: '+scenario)
-    for file, year in year_inputs.items():
-        temp = gpd.read_file('InMap/MIP_InMap_Output/'+scenario+'/'+model+'/ISRM_result_'+f"{year}"+'.shp')
-        temp['year']=year
-        temp['scenario']=scenario
-        emissions = pd.concat([emissions, temp])
-        print(year)
+for model in models:
+    for scenario in scenario_dictionary:
+        print('reading: '+scenario+ model)
+        for file, year in year_inputs.items():
+            if not os.path.exists('InMap/MIP_InMap_Output/'+scenario+'/'+model+'/ISRM_result_'+f"{year}"+'.shp'):
+                print(f"'{scenario}' emissions output doesn't exist... skipping...")
+                continue
+            temp = gpd.read_file('InMap/MIP_InMap_Output/'+scenario+'/'+model+'/ISRM_result_'+f"{year}"+'.shp')
+            temp['year']=year
+            temp['scenario']=scenario
+            temp['model']=model
+            emissions = pd.concat([emissions, temp])
+            print(year)
 
-#make file directories
-for scenario in scenario_dictionary:
-    if not os.path.exists('MIP_AirPollution/Figures/Output/' + scenario + '/' + model + '/'):
-        os.makedirs('MIP_AirPollution/Figures/Output/' + scenario + '/' + model )
+            #make file directories
+            if not os.path.exists('MIP_AirPollution/Figures/Output/' + scenario + '/' + model + '/'):
+                os.makedirs('MIP_AirPollution/Figures/Output/' + scenario + '/' + model )
 
-    if not os.path.exists('MIP_results_comparison/'+scenario+'/AirPollution/'):
-        os.makedirs('MIP_results_comparison/'+scenario+'/AirPollution')
[…]

### 2024-12-04 Downscaled/InMAP_SR.py
update to how deaths are calculated -- account for race-specific population growth

@@ -25,6 +25,8 @@ sys.path.append(os.path.abspath('C:/Users/laure/Documents/Switch-USA-PG/'))
 from MIP_AirPollution.Downscaled.settings import *
 models=['USENSYS']
 
+popgrowth = pd.read_excel('Data/np2023-t4.xlsx', sheet_name='Sheet1')
+
 def rect(i, w, s, e, n):
     x = [w[i], e[i], e[i], w[i], w[i]]
     y = [s[i], s[i], n[i], n[i], s[i]]
@@ -42,7 +44,7 @@ def poly(sr):
                             [x[3],y[3]],[x[4],y[4]]]))
     return ret
 
-def run_sr(emis, model, emis_units="tons/year"):
+def run_sr(emis, model, year, emis_units="tons/year"):
     start = time.time()
     url = 's3://inmap-model/isrm_v1.2.1.zarr/'
     fs = s3fs.S3FileSystem(anon=True, client_kwargs=dict(region_name='us-east-2'))
@@ -105,25 +107,44 @@ def run_sr(emis, model, emis_units="tons/year"):
     Native = sr['Native'][0:52411]
     Latino = sr['Latino'][0:52411]
     WhiteNoLat = sr['WhiteNoLat'][0:52411]
-    deathsK = (np.exp(np.log(1.06)/10 * TotalPM25) - 1) * TotalPop * 1.07353545 * MortalityRate / 100000 * 1.025229357798165
-    deathsL = (np.exp(np.log(1.14)/10 * TotalPM25) - 1) * TotalPop * 1.07353545 * MortalityRate / 100000 * 1.025229357798165
 
-    AsiandeathsK = (np.exp(np.log(1.06)/10 * TotalPM25) - 1) * Asian * 1.07353545 * MortalityRate / 100000 * 1.025229357798165
-    AsiandeathsL = (np.exp(np.log(1.14)/10 * TotalPM25) - 1) * Asian * 1.07353545 * MortalityRate / 100000 * 1.025229357798165
+    pop_growth_value = popgrowth.loc[popgrowth['Race'] == 'Total', year].values[0]
+    pop_growth_base = popgrowth.loc[popgrowth['Race'] == 'Total', 2010].values[0]
+    deathsK = (np.exp(np.log(1.06)/10 * TotalPM25) - 1) * TotalPop * pop_growth_value/pop_growth_base * MortalityRate / 100000
[…]

### 2024-12-04 MIP_Downscaled.tex

\documentclass[a4paper]{article}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=3cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
%% Useful packages
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{bbm}
\usepackage{setspace}
\usepackage{footmisc}
\usepackage{pdflscape}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{bm}

\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{url}

\renewcommand\footnotelayout{\fontsize{10}{12}\selectfont}

\interfootnotelinepenalty=10000

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\providecommand{\C}{\mathbb{C}}

\theoremstyle{definition}
\newtheorem{defin}{Definição}

\theoremstyle{plain}
\newtheorem{theorem}[defin]{Teorema}
\newtheorem{corollary}[defin]{Corolário}

\linespread{2}
\title{The Distributional Impacts of the Electricity Transition on Pollution Exposure}

\author{Lauren Beatty\thanks{Environmental Defense Fund  \hspace{.5cm} E-mail: lbeatty1@edf.org \hspace{.5cm}Website: \href{https://lbeatty1.github.io}{https://lbeatty1.github.io}}, 
Matthias Fripp\thanks{Environmental Defense Fund}, 
Greg Schivley, 
Rangrang Zhang, 
Cameron Wade, 
Oleg Lugovoy, 
Michael Roberts, 
Patricia Hildalgo}

\date{\today}

\begin{document}
\maketitle
\begin{center}
    PRELIMINARY DRAFT - PLEASE DO NOT CITE OR DISTRIBUTE
\end{center}

\begin{abstract}
    In this paper, we provide an overview of how different scenarios for the electricity transition will affect the distribution of air pollution exposure through 2050.  Currently, there are large disparities in pollution exposure across racial and socio-economic groups resulting from historical choices about where generation facilities were cited as well as differences across groups in income and wealth. We use results from the Model Intercomparison Project -- a project aimed at harmonizing the inputs and comparing the outputs of multiple open-source least-cost electricity planning models.  We take those results, predict location-specific emissions of various pollutants over time, then use InMap to calculate average group-level exposures to PM2.5.  We find broad agreement across models.  Moreover, we find that deep decarbonization should lead to major improvements in air quality, and reductions in pollution exposure, along with a reduction in pollution exposure disparities.
\end{abstract}

\newpage
\section{Introduction}
Historical inequities have led to large contemporary differences in air pollution exposure across groups.  This paper looks forward and analyzes how different plans for the U.S. electricity transition will affect group-level pollution exposure.

There is a rich literature on the effects of electricity generation on pollution, the effects of pollution on human health and well-being, and papers linking electricity generation to health effects.  Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acture lower respiratory illness (ALRI), cerebrovascular disease (CEV), ischaemic heart disease (IHD), and lung cancer, and in the United States, power generation is responsible for $31\%$ of premature mortality linked to outdoor air pollution \citep{Lelieveld2015TheScale}.

There is also a rich literature on how policies affect air pollution from power generation.  \citet{Shawhan2024PoliciesAmericans} use the  Engineering, Economic, and Environmental Electricity Simulation Tool (E4ST) to model capacity expansion and system operation combined with InMap to model how policies aimed at reducing air pollution in environmentally overburdened, disadvantaged communities (EO DACs) compare with hypothetical $\$15$ per ton carbon tax.  They find that both CO$_2$ pricing and the best-performing within-DAC policies reduce mortality by approximately 1.4 to 1.5 times as much for EO DAC residences as non-EO DAC residents -- a major win for reducing disparities.  However, they find that the targeted policies are only slightly more effective than the carbon tax at reducing PM2.5, and that the plans are far more costly than a carbon tax.  Similarly, \citet{Goforth2022AirStrategies} use the Regional Energy Deployment System (ReEDS) along with InMAP to project how different U.S. national decarbonization strategies will affect air pollution disparities.

This paper extends the literature by projecting air pollution impacts for a wider variety of policy scenarios, across multiple open-source least-cost electricity models.  Our results are in broad agreement with the studies of \citet{Goforth2022AirStrategies} and \citet{Shawhan2024PoliciesAmericans}.  We find that decarbonization strategies should lead to a large decrease in pollution exposure, as well as  a decline in disparities across groups in exposure.  Moreover, we find strong agreement between models.

\section{Data}

\subsection{Emissions Data}
Emissions data comes from two sources: EPA's Clean Air Markets Program Data (CAMD) and EPA's National Emissions Inventory (NEI).  The CAMD was created to track compliance with clean air programs, and thus, tracks emissions of CO$_2$, NO$_x$, and SO$_2$.  The NEI was created to track criteria pollutants, criteria precursors, and hazardous air pollutants from \textit{all} sources.  I get emissions of NH$_3$, VOC, and PM2.5 from the NEI.

\subsection{Data for PowerGenome}
Data for the MIP was generated using PowerGenome, then modified on a model-by-model basis.  PowerGenome was designed to produce input files for the GenX capacity expansion model.  PowerGenome relies on a wide variety of data such as the National Renewable Energy Lab's (NREL) Annual Technology Baseline (ATB), the U.S. Energy Information Administration's Annual Energy Outlook, the EPA's Integrated Planning Model, and more.  For more information on PowerGenome, see the release page \citep{PowerGenome/PowerGenome:V0.6.3}.  Further details on the construction of model inputs for TEMOA, SWITCH, and USENSYS can be found in the main MIP paper at INSERT PAPER HERE.

\section{Methods}

\subsection{Calculating source-specific emissions rates}
\begin{enumerate}
    \item Join EIA generation data with EPA emissions data to calculate emissions rates in lbs/MWh.
    \item Calculate emissions from existing plants by taking model-outputted generation and multiplying that by the plant's share of cluster capacity to get a predicted MWh.  Then multiply that by the plant's emissions rates for all five pollutants.
    \item New fossil generation outputted by these models is only natural gas.  For new plants, I start by allocating capacity to both retired and existing sites, with a preference for siting at retired sites.
\end{enumerate}

https://ehp.niehs.nih.gov/doi/10.1289/EHP9001
https://www.rff.org/publications/working-papers/policies-for-reducing-the-impacts-of-power-sector-air-pollution-on-disadvantaged-americans/

\subsection{Least-cost modeling}
We use four different open-source least-cost electricity planning models to solve many different scenarios. For more information on the models and scenarios, see OTHER PAPER.

\subsection{Emissions to Exposure: InMap}
Once we've obtained predictions for plant-level emissions, we feed these emissions into the InMap Source-Receptor Matrix (ISRM).  The ISRM was developed to further simply the InMAP model.  The InMAP model was designed to compute annual-average primary and secondary PM2.5.  InMAP is not a full Eulerian chemical tranport model. Rather, InMAP uses information from chemical transport models along with some simplifying assumptions and estimated reduced-form parameters \citep{Tessum2017InMAP:Interventions}.  The InMAP ISRM was developed by running InMAP hundreds of thousands of times to calculate the average linear relationship between emissions in each source location with annual-average PM2.5 concentrations at receptor locations \citep{Goodkind2019Fine-scaleEmissions}.  The InMAP ISRM uses a variable grid with a finer scale in more population-dense areas, and a coarser definition in less dense areas, greatly reducing the size of the matrix while still maintaining it's ability to make fine-scale spatial prediction.  We plot an example of the ISRM output below in figure \ref{ISRMexample}.
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figures/Output/full-current-policies/GenX/ISRM_2027_TotalPM25_concentrationmap.jpg}
        \caption{2027}
        \label{Subfig1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figures/Output/full-current-policies/GenX/ISRM_2050_TotalPM25_concentrationmap.jpg}
        \caption{2050}
        \label{Subfig2}
    \end{subfigure}\\
    \caption{ISRM Output for the Current Policies Scenario in 2027 and 2050.}
\end{figure}
We translate concentrations into mortality using total PM2.5 predicted by INMAP in a Cox proportional hazards equation:

\begin{equation}
    Deaths = [\exp(\log(r)/10*PM2.5)-1]\times Pop \times \frac{Pop_y}{Pop_{2010}} \times MortalityRate
\end{equation}
where 
\begin{itemize}
    \item $\frac{Pop_y}{Pop_{2010}}=1.0735$ is the ratio between year-$y$ population and year-2010 population (what's natively in the model).  For each model period we calculate this ratio from the 2023 U.S. Census Bureau's National Population Projection Tables.  These tables have projections by race and by year, so our estimates will reflect, for example, the fact that the Hispanic and Latino population is expected to grow by 42$\%$ by 2050, while the Non-Hispanic White population is expected to decrease \citep{Bureau2023Series}.
    \item $MortalityRate$ is the baseline mortality rate.
    \item $r$ is a parameter describing how mortality rates respond to a 10$\mu g/m^3$ increase in PM2.5 concentrations.  For this paper we use 1.06, reported by \citet{Krewski2009ExtendedInstitute}.
\end{itemize}

We project deaths forward but keep the calculation of deaths constant.  That means that our projections will not account for idiosyncratic differences across space or groups in migration patterns, or changes to the mortality rate.  Our primary concern is differences in mortality \textit{rates} rather than aggregate mortality.  Moreover, we primarily want our estimates to reflect changes in electricity generation, rather than population projections with wide uncertainty bands.
\section{Results: Comparison Across Models}

\section{Results: Comparison Across Scenarios}

Given the precise agreement between the models, we now turn our attention to comparing pollution exposure across policy scenarios.  In figure \ref{Compare Scenarios}, we plot population-weighted mean death rates by racial group by scenario for each model period.  The first thing that becomes immediately apparent is that the decarbonization scenarios reduce pollution exposure much more quickly and dramatically than the current policy scenarios.  It's also clear that there are wide disparities across racial groups, these disparities decrease in aggregate magnitudes but remain more constant in relative magnitudes over time.
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2027.jpg}
        \caption{2027}
        \label{Subfig1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2030.jpg}
        \caption{2030}
        \label{Subfig2}
    \end{subfigure}\\

        \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2030.jpg}
        \caption{2035}
        \label{Subfig2}
    \end{subfigure}
    \hfill
        \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2035.jpg}
        \caption{2040}
        \label{Subfig2}
    \end{subfigure}\\

    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2045.jpg}
        \caption{2045}
        \label{Subfig2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2050.jpg}
        \caption{2050}
        \label{Subfig2}
    \end{subfigure}
    \caption{Comparison results by scenario over time}
    \label{Compare Scenarios}
\end{figure}

FIGURE ON CO2 Vs Exposure

FIGURE ON TOTAL COST VS EXPOSURE
\subsection{Foresight}

\subsection{Constraining Transmission Expansion}

\section{Conclusion}

\begin{singlespace}
\newpage
\bibliographystyle{jpe}
%\bibliographystyle{econometrica}
\bibliography{references.bib}
\end{singlespace}

\end{document}

### 2024-12-05 Downscaled/Plot_Output_ISRM.py
lots of new figs

@@ -16,75 +16,89 @@ os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
 sys.path.append(os.path.abspath('C:/Users/laure/Documents/Switch-USA-PG/'))
 
 from MIP_AirPollution.Downscaled.settings import *
-models = ['GenX', 'SWITCH', 'TEMOA', 'USENSYS']
 
+#rather than read in all data (very time consuming)
+#I'm going to selectively plot model/scenarios
+#I'll keep them in a data dictionary so I don't need to repeatedly read them in
+data_dict = {}
 
-#%%
-## Read in all data
-emissions = pd.DataFrame()
-
-for model in models:
-    for scenario in scenario_dictionary:
-        print('reading: '+scenario+ model)
-        for file, year in year_inputs.items():
-            if not os.path.exists('InMap/MIP_InMap_Output/'+scenario+'/'+model+'/ISRM_result_'+f"{year}"+'.shp'):
-                print(f"'{scenario}' emissions output doesn't exist... skipping...")
-                continue
-            temp = gpd.read_file('InMap/MIP_InMap_Output/'+scenario+'/'+model+'/ISRM_result_'+f"{year}"+'.shp')
-            temp['year']=year
-            temp['scenario']=scenario
-            temp['model']=model
-            emissions = pd.concat([emissions, temp])
-            print(year)
-
-            #make file directories
-            if not os.path.exists('MIP_AirPollution/Figures/Output/' + scenario + '/' + model + '/'):
-                os.makedirs('MIP_AirPollution/Figures/Output/' + scenario + '/' + model )
-
-            if not os.path.exists('MIP_results_comparison/'+scenario+'/AirPollution/'):
-                os.makedirs('MIP_results_comparison/'+scenario+'/AirPollution')
-
-
-#Plot output -- map of concentrations
 #read in states
 states = gpd.read_file('Data/tl_2022_us_state/tl_2022_us_state.shp')
 states = states[~states['STUSPS'].isin(['AS', 'AK', 'GU', 'MP', 'VI', 'HI', 'PR'])]
 
-print('finding intersection with land')
-## Make all crs equal and throw out grid cells outside of U.S.
-states = states.to_crs(emissions.crs)
-intersects = emissions.geometry.intersects(states.unary_union)
-emissions['intersects'] = intersects
-emissions = emissions[emissions['intersects']==True]
+## Read in data
[…]

### 2024-12-05 MIP_Downscaled.tex

@@
 \author{Lauren Beatty\thanks{Environmental Defense Fund  \hspace{.5cm} E-mail: lbeatty1@edf.org \hspace{.5cm}Website: \href{https://lbeatty1.github.io}{https://lbeatty1.github.io}}, 
-Matthias Fripp\thanks{Environmental Defense Fund}, 
+Matthias Fripp, 
 Greg Schivley, 
 Rangrang Zhang, 
-Cameron Wade, 
+Cameron Wade, \\ 
 Oleg Lugovoy, 

### 2024-12-05 Downscaled/Plot_Output_ISRM.py
switch figure script to functions

@@ -67,252 +67,260 @@ def find_intersection_with_states(input_gdf, states_gfd):
 
 #%%
 # #Make concentration plots
+def make_concentration_plots(models, scenarios, columns, states):
+    print('making concentration plots')
+    for model in models:
+        for scenario in scenarios:
+            quant_dict= {}
+            key = f'{model}_{scenario}'
+            if key in data_dict:
+                plot_data = data_dict[key]
+            else:
+                data_dict[key]=read_scenario_data(scenario, model, states)
+                plot_data = data_dict[key]
+
+            for file,year in year_inputs.items():
+                plot_temp = plot_data[plot_data['year']==year]
+                for column in columns_to_plot:
+
+                    #want scales to be the same for each plot
+                    if year==2027:
+                        print("Defining scale")
+                        q = 0.99  # Truncate results at the 99th percentile for better visualization
+                        cut = np.quantile(plot_temp[column], q)
+                        quant_dict[column]=cut
+
+                    fig, ax = plt.subplots(figsize=(10, 8))
+
+                    # Plot data with color scale truncated at the specified quantile
+                    plot_temp.plot(vmin=0, vmax=cut, cmap="OrRd", column=column, ax=ax)
+
+                    # Plot states boundaries
+                    states = states.to_crs(plot_temp.crs)
+                    states.boundary.plot(ax=ax, color='black', linewidth=0.2)
+
+                    # Add color bar
+                    sm = plt.cm.ScalarMappable(cmap='OrRd', norm=plt.Normalize(vmin=0, vmax=quant_dict[column]))
+                    sm._A = []  # Fake empty array for the colorbar
+                    cbar = fig.colorbar(sm, fraction=0.03)  
+                    cbar.set_label(column+' concentration (μg m$^{-3}$)', fontsize=18)  # Set color bar label
+                    cbar.ax.tick_params(labelsize=14)
+
+                    plt.axis('off')
+                    plt.savefig(f'MIP_AirPollution/Figures/Output/{scenario}/{model}/ISRM_{year}_{column}_concentrationmap.jpg', format='jpg',
+                                dpi=300, bbox_inches='tight')
[…]

### 2024-12-05 MIP_Downscaled.tex
Updates from Overleaf

@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figures/Output/full-current-policies/GenX/ISRM_2027_TotalPM25_concentrationmap.jpg}
+        \includegraphics[width=\textwidth]{Figures/Output/full-base-200/GenX/ISRM_2027_TotalPM25_concentrationmap.jpg}
         \caption{2027}
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figures/Output/full-current-policies/GenX/ISRM_2050_TotalPM25_concentrationmap.jpg}
+        \includegraphics[width=\textwidth]{Figures/Output/full-base-200/GenX/ISRM_2050_TotalPM25_concentrationmap.jpg}
         \caption{2050}
@@
     \end{subfigure}\\
-    \caption{ISRM Output for the Current Policies Scenario in 2027 and 2050.}
+    \caption{ISRM Output for the Base Scenario in 2027 and 2050.}
 \end{figure}

### 2024-12-05 Downscaled/Plot_Output_ISRM.py
fix lines in plot

@@ -26,6 +26,16 @@ data_dict = {}
 states = gpd.read_file('Data/tl_2022_us_state/tl_2022_us_state.shp')
 states = states[~states['STUSPS'].isin(['AS', 'AK', 'GU', 'MP', 'VI', 'HI', 'PR'])]
 
+#keep palette consistent across plots
+palette = sns.color_palette("bright")
+color_mapping = {
+    'Latino': palette[0],       # Red
+    'WhiteNoL_1': palette[1],  # Blue
+    'Black': palette[2],       # Green
+    'Asian': palette[3],       # Purple
+    'Native': palette[4]       # Orange
+}
+
 ## Read in data
 def read_scenario_data(scenario, model, states_gdf):
     result = pd.DataFrame()
@@ -139,10 +149,10 @@ def make_death_timeseries(models, scenarios):
             print('deaths: '+scenario)
 
             # Plot mortality rates
-            plt.plot(deaths['year'], deaths['Latinodeat'], marker='o', label='Latino Deaths')
-            plt.plot(deaths['year'], deaths['WhiteNoL_1'], marker='o', label='White/Non-hispanic Deaths')
-            plt.plot(deaths['year'], deaths['Blackdeath'], marker='o', label='Black Deaths')
-            plt.plot(deaths['year'], deaths['Asiandeath'], marker='o', label='Asian Deaths')
+            plt.plot(deaths['year'], deaths['Latinodeat'], marker='o', label='Latino Deaths', color=color_mapping['Latino'])
+            plt.plot(deaths['year'], deaths['WhiteNoL_1'], marker='o', label='White/Non-hispanic Deaths', color=color_mapping['WhiteNoL_1'])
+            plt.plot(deaths['year'], deaths['Blackdeath'], marker='o', label='Black Deaths', color=color_mapping['Black'])
+            plt.plot(deaths['year'], deaths['Asiandeath'], marker='o', label='Asian Deaths', color=color_mapping['Asian'])
 
             # Add labels and title
             plt.xlabel('Year', fontsize=16)
@@ -385,6 +395,7 @@ for scenario in scenarios:
                     dpi=300, bbox_inches='tight')
 
 
+
 #%%
 #######################################
 #### COMPARE WITH CO2 EMISSIONS #######
@@ -418,7 +429,7 @@ for model in models:
 
[…]

### 2024-12-05 MIP_Downscaled.tex

@@
 \end{figure}
-We translate concentrations into mortality using total PM2.5 predicted by INMAP in a Cox proportional hazards equation:
+For each grid area, $i$, we translate concentrations into mortality using total PM2.5 predicted by INMAP in a Cox proportional hazards equation:
 
 \begin{equation}
-    Deaths = [\exp(\log(r)/10*PM2.5)-1]\times Pop \times \frac{Pop_y}{Pop_{2010}} \times MortalityRate
+    Deaths_i = [\exp(\log(r)/10*PM2.5_i)-1]\times Pop_i \times \frac{Pop_y}{Pop_{2010}} \times MortalityRate_i
 \end{equation}
@@
 \begin{itemize}
-    \item $\frac{Pop_y}{Pop_{2010}}=1.0735$ is the ratio between year-$y$ population and year-2010 population (what's natively in the model).
+    \item $r$ is a parameter describing how mortality rates respond to a 10$\mu g/m^3$ increase in PM2.5 concentrations.  For this paper we use 1.06, reported by \citet{Krewski2009ExtendedInstitute}.
+    \item $Pop_i$ is the population of grid $i$ in 2010.
+    \item $\frac{Pop_y}{Pop_{2010}}$ is the ratio between year-$y$ population and year-2010 population (what's natively in the model).
 For each model period we calculate this ratio from the 2023 U.S.
@@
 These tables have projections by race and by year, so our estimates will reflect, for example, the fact that the Hispanic and Latino population is expected to grow by 42$\%$ by 2050, while the Non-Hispanic White population is expected to decrease \citep{Bureau2023Series}.
-    \item $MortalityRate$ is the baseline mortality rate.
-    \item $r$ is a parameter describing how mortality rates respond to a 10$\mu g/m^3$ increase in PM2.5 concentrations.  For this paper we use 1.06, reported by \citet{Krewski2009ExtendedInstitute}.
+    \item $MortalityRate_i$ is the mortality rate for grid $i$.\footnote{The mortality rate in the model is for 2005.
+For simplicty, we don't replace this.
+Since it remains constant over time, these projections will not account for changes in mortality rates over time.}
 \end{itemize}
 
-We project deaths forward but keep the calculation of deaths constant.
-That means that our projections will not account for idiosyncratic differences across space or groups in migration patterns, or changes to the mortality rate.
-Our primary concern is differences in mortality \textit{rates} rather than aggregate mortality.
+While the projection takes into account idiosyncratic changes across groups in population growth, it will not take into acount idiosyncratic changes across groups \textit{across space}.
+In addition, our estimates will not reflect differences between groups in mortality rates, which previous work has shown may be an important driver of mortality in models \citep{Spiller2021MortalityOutcomes}.
+Lastly, these projections will not take into account expected changes over time in mortality rates.
+Since our primary concern is differences in mortality \textit{rates} rather than aggregate mortality, we think these are reasonable assumptions and simplifications.
 Moreover, we primarily want our estimates to reflect changes in electricity generation, rather than population projections with wide uncertainty bands.
 \section{Results: Comparison Across Models}
+Our first result is that there is tight agreement across models.
+This is a fairly unsurprising result in light of the results of MIP PROJECT PAPER.
+The models have similar amounts of generation and capacity expansion across models, as well as similar CO$_2$ emissions and costs.
+The largest difference we observe between race-specific death rates is X$\%$.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figures/Output/DeathRate_Across_Models_current_policies_52_week_2027.png}
+        \label{2027CompareModels}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figures/Output/DeathRate_Across_Models_current_policies_52_week_2050.png}
+        \label{2050CompareModels}
+    \end{subfigure}\\
+
+    \caption{Comparison of results across models for the current policies scenario in 2027 and 2050}
+    \label{Compare Scenarios}
+\end{figure}
 
@@
     \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2027.jpg}
+    \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{Figures/Output/Compare_scenarios_all-scenarios_2027_GenX.jpg}
         \caption{2027}
@@
     \hfill
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2030.jpg}
+    \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{Figures/Output/Compare_scenarios_all-scenarios_2030_GenX.jpg}
         \caption{2030}
@@
 
-        \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2030.jpg}
+        \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{Figures/Output/Compare_scenarios_all-scenarios_2035_GenX.jpg}
         \caption{2035}
@@
     \hfill
-        \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2035.jpg}
+        \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{Figures/Output/Compare_scenarios_all-scenarios_2040_GenX.jpg}
         \caption{2040}
@@
 
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2045.jpg}
+    \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{Figures/Output/Compare_scenarios_all-scenarios_2045_GenX.jpg}
         \caption{2045}
@@
     \hfill
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2050.jpg}
+    \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{Figures/Output/Compare_scenarios_all-scenarios_2050_GenX.jpg}
         \caption{2050}
@@
 
-FIGURE ON CO2 Vs Exposure
+In figure \ref{DeathsCosts} we plot deaths against total system costs.
+The policies that seemingly do the best in terms of cost-effectiveness are the full-base-200 scenarios.
+NOTE FOR DISCUSSION -- THIS IS ALMOST BY CONSTRUCTION SINCE CO2 COSTS ARE INCLUDED IN TOTAL COSTS
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{Figures/Output/Deaths_vs_Costs_GenX.png}
+    \caption{Plot of total system costs versus average deaths per year.  Results are from GenX output.}
+    \label{DeathsCosts}
+\end{figure}
 
-FIGURE ON TOTAL COST VS EXPOSURE
-\subsection{Foresight}
+Unsurprisingly, deaths are highly correlated with CO$_2$ emissions, since both are driven by fossil-fuel combustion.
+Policies that most dramatically reduce CO$_2$ emissions, also do the most to reduce aggregate deaths.
+\begin{figure}
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/Output/Deaths_vs_Emissions_GenX.png}
+    \caption{Deaths are highly correlated with CO$_2$ emissions}
+    \label{DeathsEmissions}
+\end{figure}
+
+\subsection{Varying the Escape Price}
+In figure \ref{EscapePrice} we plot death rates over time across different escape-price scenarios.
+The difference between a $\$50$ escape price and a $\$200$ escape price is quite striking.
+However, the change from $\$200$ to $\$1,000$ is far more subtle, and mostly shows up in 2040 and beyond.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/Output/Compare_scenarios_compare-escape-price_GenX.jpg}
+    \caption{Death rates over time across escape-price scenarios}
+    \label{EscapePrice}
+\end{figure}
 
 \subsection{Constraining Transmission Expansion}
+In figure \ref{TransmissionConstraint} we plot death rates over time across the transmission constraint scenarios.
+Transmission constraints appear to have almost no effect on deaths.
+As explored in MIP PAPER, the effects of transmission constraints are mostly localized.
+\begin{figure}
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/Output/Compare_scenarios_transmissions-constraint_GenX.jpg}
+    \caption{Death rates over time for transmission constraint scenarios}
+    \label{TransmissionConstraint}
+\end{figure}
 
 \section{Conclusion}
+This paper presents projections for how numerous different transition scenarios calculated using four different open-source models will affect the distribution of pollution exposure through 2050.
+We find that pollution projections are highly synchronized across models.
+Moreover, we find that a more aggressive energy transition could save more than 2,000 lives per year.
 

### 2024-12-05 MIP_Downscaled.tex

@@
     \centering
-    \includegraphics[width=1\linewidth]{Figures/Output/Deaths_vs_Emissions_GenX.png}
-    \caption{Deaths are highly correlated with CO$_2$ emissions}
+    \includegraphics[width=1\linewidth]{Figures/Output/Deaths_vs_Emissions_withlines_GenX.png}
+    \caption{Deaths are highly correlated with CO$_2$ emissions.
+In this plot, points are colored by scenarios and shaped by model year.
+We also draw lines to show the trajectory over time for each model.}
     \label{DeathsEmissions}

### 2024-12-09 Downscaled/Plot_Output_ISRM.py
new cost plot

@@ -572,3 +572,72 @@ plt.savefig(f'MIP_AirPollution/Figures/Output/Deaths_vs_Costs_{model}.png', form
             dpi=300, bbox_inches='tight')
 
 # %%
+#########################
+### Break down costs
+#########################
+model='GenX'
+scenarios = ['full-base-50', 'full-base-200', 'full-base-1000','full-base-200-tx-0', 'full-base-200-tx-15', 'full-base-200-tx50', 'full-current-policies', 'full-current-policies-retire', 'full-current-policies-commit']
+costs = pd.read_csv('MIP_results_comparison/compiled_results/all/operational_costs.csv')
+costs = costs[costs['planning_year']=='NPV']
+costs = costs[costs['case'].isin(scenarios)]
+costs = costs[costs['model']==model]
+#read in data
+scenarios_compare = pd.DataFrame()
+scenarios_name = '_'.join(scenarios)
+for scenario in scenarios:
+    key = f'{model}_{scenario}'
+    if key in data_dict:
+        plot_data = data_dict[key]
+        print(f'{key} already in memory')
+    else:
+        data_dict[key]=read_scenario_data(scenario, model, states)
+        plot_data = data_dict[key]
+    scenarios_compare = pd.concat([scenarios_compare, plot_data])
+scenarios_compare = scenarios_compare.groupby(['scenario', 'model', 'year']).sum().reset_index()
+    
+scenarios_compare['Total']=scenarios_compare['deathsK']*7.4*1e6 #VSL
+scenarios_compare['Costs'] = 'Deaths'
+scenarios_compare['case']=scenarios_compare['scenario']
+scenarios_compare['Total'] = np.where(scenarios_compare['year'] == 2027, scenarios_compare['Total'] * 3, scenarios_compare['Total'] * 5)
+scenarios_compare = scenarios_compare.groupby(['Costs', 'case']).agg({'Total':'sum'}).reset_index()
+costs = pd.concat([costs, scenarios_compare])
+
+costs = costs[['Costs', 'Total', 'case']]
+costs['Total']=costs['Total']/1e9
+
+
+#simplify and control order of bars
+other_costs = ['cNSE', 'cNetworkExp', 'cStart', 'cUnmetRsv', 'cVOM']
+costs['Costs'] = np.where(costs['Costs'].isin(other_costs), 'Other', costs['Costs'])
+
[…]

### 2024-12-09 MIP_Downscaled.tex
Updates from Overleaf

@@
     In this paper, we provide an overview of how different scenarios for the electricity transition will affect the distribution of air pollution exposure through 2050.
-Currently, there are large disparities in pollution exposure across racial and socio-economic groups resulting from historical choices about where generation facilities were cited as well as differences across groups in income and wealth.
+Currently, there are large disparities in pollution exposure across racial and socioeconomic groups resulting from historical choices about where generation facilities were cited, historical policies affecting migration and housing choices, and differences across groups in income and wealth.
 We use results from the Model Intercomparison Project -- a project aimed at harmonizing the inputs and comparing the outputs of multiple open-source least-cost electricity planning models.
@@
 Historical inequities have led to large contemporary differences in air pollution exposure across groups.
+The academic and policy communities have increasingly been turning their attention towards reducing these inequities, and analyzing the equity impacts of policy.
 This paper looks forward and analyzes how different plans for the U.S.
@@
 
-There is a rich literature on the effects of electricity generation on pollution, the effects of pollution on human health and well-being, and papers linking electricity generation to health effects.
+There is a rich literature on the effects of electricity generation on pollution and the effects of pollution on human health and well-being.
 Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acture lower respiratory illness (ALRI), cerebrovascular disease (CEV), ischaemic heart disease (IHD), and lung cancer, and in the United States, power generation is responsible for $31\%$ of premature mortality linked to outdoor air pollution \citep{Lelieveld2015TheScale}.
@@
 \citet{Shawhan2024PoliciesAmericans} use the  Engineering, Economic, and Environmental Electricity Simulation Tool (E4ST) to model capacity expansion and system operation combined with InMap to model how policies aimed at reducing air pollution in environmentally overburdened, disadvantaged communities (EO DACs) compare with hypothetical $\$15$ per ton carbon tax.
-They find that both CO$_2$ pricing and the best-performing within-DAC policies reduce mortality by approximately 1.4 to 1.5 times as much for EO DAC residences as non-EO DAC residents -- a major win for reducing disparities.
-However, they find that the targeted policies are only slightly more effective than the carbon tax at reducing PM2.5, and that the plans are far more costly than a carbon tax.
+They find that both CO$_2$ pricing and the best performing within-DAC policies reduce mortality by approximately 1.4 to 1.5 times as much for EO DAC residences as non-EO DAC residents, a major win for reducing disparities.
+However, they find that targeted policies are only slightly more effective than the carbon tax at reducing PM2.5, and that the plans are far more costly than a carbon tax.
 Similarly, \citet{Goforth2022AirStrategies} use the Regional Energy Deployment System (ReEDS) along with InMAP to project how different U.S.
@@
 \section{Methods}
+This paper accompanies the Model Intercomparison Project.
+Once MIP outputs were obtained, we calculate source-specific emissions rates using empirical emissions data, downscale the results from MIP to predict location-specific aggregate emissions, then finally, run those emissions through the InMAP ISRM model.
+
+\subsection{Least-cost modeling}
+We use four different open-source least-cost electricity planning models to solve many different scenarios. For more information on the models and scenarios, see OTHER PAPER.
 
@@
 \begin{enumerate}
-    \item Join EIA generation data with EPA emissions data to calculate emissions rates in lbs/MWh.
+    \item Join EIA generation data with EPA emissions data to calculate emissions rates in tons/MWh.
+Many plants are in both the EPA and EIA data and can be matched directly, and have their own specific emissions rates.
+For plants that cannot be matched, we take generation-weighted average emissions rates for each technology.
     \item Calculate emissions from existing plants by taking model-outputted generation and multiplying that by the plant's share of cluster capacity to get a predicted MWh.
 Then multiply that by the plant's emissions rates for all five pollutants.
-    \item New fossil generation outputted by these models is only natural gas.
-For new plants, I start by allocating capacity to both retired and existing sites, with a preference for siting at retired sites.
+    \item The new fossil generation produced by these models is only natural gas or natural gas with CCS.
+For new plants, we begin by allocating new capacity to retired sites.
+Once the new capacity equals the retired capacity, we allocate the remaining new capacity to both existing and retired sites, weighted by capacity.
+This is essentially an assumption that the spatial distribution of new generation will mirror existing and retired generation.
 \end{enumerate}
-
-https://ehp.niehs.nih.gov/doi/10.1289/EHP9001
-https://www.rff.org/publications/working-papers/policies-for-reducing-the-impacts-of-power-sector-air-pollution-on-disadvantaged-americans/
-
-\subsection{Least-cost modeling}
-We use four different open-source least-cost electricity planning models to solve many different scenarios. For more information on the models and scenarios, see OTHER PAPER.
 
@@
 
+These results mirror previous related work.  \citet{Shawhan2024PoliciesAmericans} found that a carbon tax would decrease deaths by X$\%$.  \citet{Goforth2022AirStrategies} find that THINGS.
+
 \subsection{Varying the Escape Price}
@@
 As explored in MIP PAPER, the effects of transmission constraints are mostly localized.
+MIP PAPER found that while the transmission constraint did not have a major effect on system costs, it did increase aggregate emissions by $X\%$.
+The effects on aggregate deaths are far more muted.
+The most severe transmission constraint (zero new inter-regional transmission) only has X more deaths per year than the completely unconstrained scenario.
 \begin{figure}
@@
 We find that pollution projections are highly synchronized across models.
-Moreover, we find that a more aggressive energy transition could save more than 2,000 lives per year.
+Moreover, we find that a more aggressive energy transition could save more than 2,000 lives per year relative to the current policies trajectory.
+All model results point towards a dramatic decrease in deaths for all racial groups.
 

### 2024-12-09 MIP_Downscaled.tex

@@
 
+The cost of deaths can be a sizable portion of overall system costs.
+In figure \ref{CostsBarchart}, we plot total costs (in NPV) across scenarios, including a component for monetized costs of air pollution deaths.
+We use EPA's Value of Statistical Life, which is currently $\$7.4 million$ in U.S.
+$\$ 2006$, which we inflation-adjust to $\$11.51$ million in 2024 dollars.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{Figures/Output/Costs_Barchart_GenX.png}
+    \caption{Caption}
+    \label{CostsBarchart}
+\end{figure}
+
 These results mirror previous related work.  \citet{Shawhan2024PoliciesAmericans} found that a carbon tax would decrease deaths by X$\%$.  \citet{Goforth2022AirStrategies} find that THINGS.

### 2024-12-09 Downscaled/Plot_Output_ISRM.py
fix fig with lines

@@ -495,7 +495,7 @@ sns.scatterplot(
 )
 
 # Add labels and title
-plt.title('Tracking Changes in X and Y Over Time', fontsize=16)
+plt.title('CO2 Emissions vs Deaths', fontsize=16)
 plt.xlabel('CO2 Emissions (100 million Tons)', fontsize=14)
 plt.ylabel('Deaths per year', fontsize=14)
 
@@ -515,6 +515,8 @@ plt.savefig(f'MIP_AirPollution/Figures/Output/Deaths_vs_Emissions_withlines_{mod
 plt.savefig(f'MIP_AirPollution/Figures/Output/Deaths_vs_Emissions_withlines_{model}.png', format='jpg',
             dpi=300, bbox_inches='tight')
 plt.close()
+
+#%%
 #######################################
 ###### COMPARE WITH COSTS #############
 #######################################
@@ -595,7 +597,7 @@ for scenario in scenarios:
     scenarios_compare = pd.concat([scenarios_compare, plot_data])
 scenarios_compare = scenarios_compare.groupby(['scenario', 'model', 'year']).sum().reset_index()
     
-scenarios_compare['Total']=scenarios_compare['deathsK']*7.4*1e6 #VSL
+scenarios_compare['Total']=scenarios_compare['deathsK']*11.51*1e6 #VSL
 scenarios_compare['Costs'] = 'Deaths'
 scenarios_compare['case']=scenarios_compare['scenario']
 scenarios_compare['Total'] = np.where(scenarios_compare['year'] == 2027, scenarios_compare['Total'] * 3, scenarios_compare['Total'] * 5)

### 2024-12-09 Downscaled/Plot_Output_ISRM.py
writing summary stats to csv

@@ -10,12 +10,19 @@ import math
 from matplotlib.colors import LinearSegmentedColormap
 import seaborn as sns
 import sys
+import csv
 
 
 os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
 sys.path.append(os.path.abspath('C:/Users/laure/Documents/Switch-USA-PG/'))
 
 from MIP_AirPollution.Downscaled.settings import *
+
+#write an empty csv to save some summary stats to
+stats_file = 'MIP_AirPollution/Downscaled/stats.csv'
+with open(stats_file, mode='w', newline='') as file:
+    writer = csv.writer(file)
+    writer.writerow([])
 
 #rather than read in all data (very time consuming)
 #I'm going to selectively plot model/scenarios
@@ -121,12 +128,24 @@ def make_concentration_plots(models, scenarios, columns, states):
                                 dpi=300, bbox_inches='tight')
                     plt.savefig(f'MIP_results_comparison/AirPollution/{scenario}/{model}/ISRM_{year}_{column}_{model}_concentrationmap.jpg', format='jpg',
                                 dpi=300, bbox_inches='tight')
+                    
+                    #save population-weighted to stats.csv
+                    plot_temp['TotalPop'] = plot_temp['Black']+plot_temp['Asian']+plot_temp['WhiteNoLat']+plot_temp['Native']+plot_temp['Latino']
+                    weighted_avg = (plot_temp['TotalPM25'] * plot_temp['TotalPop']).sum() / plot_temp['TotalPop'].sum()
+                    with open(stats_file, mode='a', newline='') as file:
+                        writer = csv.writer(file)
+                        writer.writerow([scenario, model, year, weighted_avg])
+
 
 
 columns_to_plot = ['TotalPM25']
 models = ['GenX', 'SWITCH']
 scenarios = ['full-base-200', 'full-base-50', 'full-base-1000']
-make_concentration_plots(models = models, scenarios=scenarios, states=states)
+with open(stats_file, mode='a', newline='') as file:
+    writer = csv.writer(file)
+    writer.writerow(['Population-weighted average concentrations'])
+
+make_concentration_plots(models = models, scenarios=scenarios, columns = columns_to_plot, states=states)
 
 #%%
 def make_death_timeseries(models, scenarios):
[…]

### 2025-03-06 MIP_Downscaled.tex

--- 2024-12-09 MIP_Downscaled.tex
@@
 
-\author{Lauren Beatty\thanks{Environmental Defense Fund  \hspace{.5cm} E-mail: lbeatty1@edf.org \hspace{.5cm}Website: \href{https://lbeatty1.github.io}{https://lbeatty1.github.io}}, 
-Matthias Fripp, 
-Greg Schivley, 
-Rangrang Zhang, 
-Cameron Wade, \\ 
+\author{Lauren Beatty\thanks{Environmental Defense Fund}, 
+Matthias Fripp\footnotemark[1], 
+Greg Schivley \thanks{Princeton University}, 
+Rangrang Zhang \thanks{University of Hawaii}, 
+Cameron Wade \thanks{Sutubra}, \\ 
 Oleg Lugovoy, 
-Michael Roberts, 
-Patricia Hildalgo}
+Michael Roberts\footnotemark[3], 
+Patricia Hildalgo\thanks{University of California, San Diego}}
 
@@
 Historical inequities have led to large contemporary differences in air pollution exposure across groups.
-The academic and policy communities have increasingly been turning their attention towards reducing these inequities, and analyzing the equity impacts of policy.
+The academic and policy communities have increasingly been turning their attention towards reducing these inequities, as well as analyzing the equity impacts of policy.
 This paper looks forward and analyzes how different plans for the U.S.
@@
 Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acture lower respiratory illness (ALRI), cerebrovascular disease (CEV), ischaemic heart disease (IHD), and lung cancer, and in the United States, power generation is responsible for $31\%$ of premature mortality linked to outdoor air pollution \citep{Lelieveld2015TheScale}.
+There are also many retrospective policy analysis of how policies such as the Clean Air Act, the Acid Rain Program, or even EV Tax credits affected air pollution and mortality (\citet{Cropper2023The1970s}, \citet{Chay2003TheJSTOR}, \citet{Holland2019DistributionalAdoption}, \citet{Chan2018}).
+These studies broadly find that the monetized health effects from policy changes that affect electricity generation are important components of the overall costs and benefits of these policies.
 
-There is also a rich literature on how policies affect air pollution from power generation.
+There is also an emerging literature aiming to estimate how various transition policies will affect air pollution moving forward.
 \citet{Shawhan2024PoliciesAmericans} use the  Engineering, Economic, and Environmental Electricity Simulation Tool (E4ST) to model capacity expansion and system operation combined with InMap to model how policies aimed at reducing air pollution in environmentally overburdened, disadvantaged communities (EO DACs) compare with hypothetical $\$15$ per ton carbon tax.
@@
 national decarbonization strategies will affect air pollution disparities.
+Lastly, \citet{Jordan2024ClimateGeneration} uses the multi-sectoral ability of TEMOA to jointly model the air pollution effects of model electricity generation and transportation.
 
@@
 Energy Information Administration's Annual Energy Outlook, the EPA's Integrated Planning Model, and more.
-For more information on PowerGenome, see the release page \citep{PowerGenome/PowerGenome:V0.6.3}.
+For more information on PowerGenome, see the release page \citep{SchivleyPowerGenome/PowerGenome:V0.6.3}.
 Further details on the construction of model inputs for TEMOA, SWITCH, and USENSYS can be found in the main MIP paper at INSERT PAPER HERE.
@@
 \subsection{Least-cost modeling}
-We use four different open-source least-cost electricity planning models to solve many different scenarios. For more information on the models and scenarios, see OTHER PAPER.
+We use four different open-source least-cost electricity planning models to solve many different scenarios.
+Each model is a least-cost model, so the model takes as inputs assumptions about costs for various generation technologies, assumptions about policies, assumptions about demand, and the current state of the electrical system, then aims to find a portfolio of new capacity combined with feasible hourly generation that minimizes costs subject to reliability and physical constraints.
+
+Each of these models was solved using 26 zones, and a whole year of hourly demand and weather data.  For more information on the models and scenarios, see OTHER PAPER.
 
 \subsection{Calculating source-specific emissions rates}
+The exercise here is two-fold.
+First, we need to calculate emissions rates for plants.
+Then we need to calculate model-zone specific emissions rates that are spatially-explicity.
+For example, a model run may tell you that combined-cycle natural gas in PJM-East should generate 100,000 MWh per year.
+We need to make assumptions to assign those 100,000 MWh to specific locations to calculate location-specific aggregate emissions.
+
 \begin{enumerate}
@@
 This is essentially an assumption that the spatial distribution of new generation will mirror existing and retired generation.
+In the literature, this is known as a ``grow-in-place" assumption, and is used by other studies such as \citet{Jordan2024ClimateGeneration}.
+We don't have data on emissions rates of CCS, so we inflate emissions rates for CCS by the model assumption about the ratio of CCS heat-rates to heat-rates at plants without pollution control technologies.
+Of course, if CCS technologies also scrub some or many air pollutants, our estimates of pollution will be overstated, especially in later planning years.
 \end{enumerate}
@@
     \caption{ISRM Output for the Base Scenario in 2027 and 2050.}
+    \label{ISRMexample}
 \end{figure}
@@
 Lastly, these projections will not take into account expected changes over time in mortality rates.
-Since our primary concern is differences in mortality \textit{rates} rather than aggregate mortality, we think these are reasonable assumptions and simplifications.
+If mortality rates fall over time, then our estimates will be biased upwards.
+
+Since our primary concern is differences in mortality rates across groups, rather than aggregate mortality, we think these are reasonable assumptions and simplifications.
 Moreover, we primarily want our estimates to reflect changes in electricity generation, rather than population projections with wide uncertainty bands.
+
 \section{Results: Comparison Across Models}
@@
 The models have similar amounts of generation and capacity expansion across models, as well as similar CO$_2$ emissions and costs.
-The largest difference we observe between race-specific death rates is X$\%$.
+The largest difference we observe between race-specific death rates is the 15$\%$ discrepancy between USENSYS and SWITCH in the projected death rate for Black individuals in 2050.
+For the earlier planning periods (2027 and 2030) where death rates are higher, model discrepancies are less than 10$\%$.
 
@@
 Given the precise agreement between the models, we now turn our attention to comparing pollution exposure across policy scenarios.
+For brevity, we present outputs from GenX model runs, but additional results can be found in our project dashboard.
+
 In figure \ref{Compare Scenarios}, we plot population-weighted mean death rates by racial group by scenario for each model period.
-The first thing that becomes immediately apparent is that the decarbonization scenarios reduce pollution exposure much more quickly and dramatically than the current policy scenarios.
-It's also clear that there are wide disparities across racial groups, these disparities decrease in aggregate magnitudes but remain more constant in relative magnitudes over time.
+The first thing that becomes immediately apparent is that aggressive decarbonization scenarios reduce pollution exposure much more quickly and dramatically than current policy scenarios.
+It is also clear that there are wide disparities across racial groups and that these disparities decrease in aggregate magnitudes but remain more constant in relative magnitudes over time.
+For example, in the base $\$200$ scenario, the disparity between Black and white mortality rates is 2.02 per million in 2027, or a $22\%$ difference.
+By 2050, the disparity falls to 2.63 per 10 million but this represents a 33$\%$ difference.
+
 \begin{figure}
@@
     \includegraphics[width=0.8\linewidth]{Figures/Output/Deaths_vs_Costs_GenX.png}
-    \caption{Plot of total system costs versus average deaths per year.  Results are from GenX output.}
+    \caption{Plot of total system costs versus average deaths per year.  Results are from GenX output.  Note that the monetized damages from mortality are \textit{not} included in the costs.}
     \label{DeathsCosts}
@@
 
-Unsurprisingly, deaths are highly correlated with CO$_2$ emissions, since both are driven by fossil-fuel combustion.
+In figure \ref{CostsBarchart}, we monetize the costs of mortality and show the breakdown of costs across scenarios.
+The cost of deaths can be a sizable portion of overall system costs.
+We use EPA's Value of Statistical Life, which is currently $\$7.4$ million in U.S.
+$\$ 2006$, which we inflation-adjust to $\$11.51$ million in 2024 dollars \citep{U.S.EnvironmentalProtectionAgencyMortalityEPA}.
+In the current policy scenarios, which have higher levels of emissions and deaths, the NPV of deaths is on a similar order of magnitude as total variable costs.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{Figures/Output/Costs_Barchart_GenX.png}
+    \caption{Breakdown of costs by scenario}
+    \label{CostsBarchart}
+\end{figure}
+
+Unsurprisingly, deaths are highly correlated with CO$_2$ emissions since both are driven by fossil-fuel combustion.
 Policies that most dramatically reduce CO$_2$ emissions, also do the most to reduce aggregate deaths.
+In figure \ref{DeathsEmissions} we plot emissions and deaths in various scenarios.
+This plot shows both the strong correlation between CO$_2$ emissions and shows how each scenario evolves over the model time horizon.
+
+Although this figure \textit{ does} highlight that deaths and emissions are highly correlated, it does show that the correlation is imperfect.
+For example, between 2040 and 2045, aggregate CO$_2$ emissions increase in the current policy scenarios while deaths fall.
+
 \begin{figure}
@@
 
-These results mirror previous related work.  \citet{Shawhan2024PoliciesAmericans} found that a carbon tax would decrease deaths by X$\%$.  \citet{Goforth2022AirStrategies} find that THINGS.
+These results mirror previous related work, though results are difficult to compare since each paper uses their own scenarios. 
+ \citet{Shawhan2024PoliciesAmericans} found that a $\$15$ carbon tax would decrease annual deaths by around 2,000 through 2030.
+Our estimate for the base $\$50$ scenario is that annual deaths would decrease by around 3,000 through 2030.
 
-\subsection{Varying the Escape Price}
+\subsection{Varying the Carbon/Escape Price}
 In figure \ref{EscapePrice} we plot death rates over time across different escape-price scenarios.
@@
 However, the change from $\$200$ to $\$1,000$ is far more subtle, and mostly shows up in 2040 and beyond.
+Like, CO$_2$ emissions, it appears that marginal costs of abatement (or maginal cost of reducing deaths) increase with the abatement level.
 
@@
 This paper presents projections for how numerous different transition scenarios calculated using four different open-source models will affect the distribution of pollution exposure through 2050.
-We find that pollution projections are highly synchronized across models.
+We find that the pollution projections are highly synchronized between the models.
+
 Moreover, we find that a more aggressive energy transition could save more than 2,000 lives per year relative to the current policies trajectory.
-All model results point towards a dramatic decrease in deaths for all racial groups.
+All model results point towards a dramatic decrease in deaths across all racial groups, and a convergence over time across groups in death rates.
 

### 2025-05-23 Endogenous/Endogenous_makeplots.py
results for Westerns

@@ -8,14 +8,20 @@ import pandas as pd
 import math
 from matplotlib.colors import LinearSegmentedColormap
 from matplotlib.colors import ListedColormap
+import yaml
 
 
 
 os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
 
-scenario = 'current_policies_short_simplified'
-years = [2027, 2030]
-result_dir = f'switch/26-zone/out/foresight/'
+with open("MIP_results_comparison/case_settings/26-zone/settings-atb2023/model_definition.yml", "r") as file:
+    region_aggregations = yaml.safe_load(file)["region_aggregations"]
+region_aggregations['FRCC']=['FRCC']
+
+
+scenario = 'current_policies_short'
+years = [2027, 2030, 2035, 2040, 2045, 2050]
+result_dir = f'switch/26-zone/out/'
 cost_limits = [50,100,200]
 
 if not os.path.exists('MIP_AirPollution/Figures/EndogenousResults/'+scenario+'/'):
@@ -26,7 +32,7 @@ coefs=pd.DataFrame()
 for y in years:
     coefs = pd.concat([coefs,
                        pd.read_csv(
-                           f"MIP_Air_OutData/Marginal_Coefficients/{scenario}_marginal_exposure_coefs_{y}.csv",
+                           f"MIP_Air_OutData/{scenario}_marginal_exposure_coefs_{y}.csv",
                            index_col=0)]
     )
 
@@ -55,7 +61,7 @@ print(summary)
 
 emis = []
 for year in years:
-    shapefile_path = f'MIP_Air_OutData/MIP_Emissions/{scenario}_marginal_emissions_{y}.shp'
+    shapefile_path = f'MIP_Air_OutData/{scenario}_simplified_marginal_emissions_{y}.shp'
 
     gdf = gpd.read_file(shapefile_path)
     
@@ -144,23 +150,29 @@ for c in cost_limits:
     tempdat['scenario']=f'{scenario}_{c}'
     capacity = pd.concat([capacity,tempdat])
 
-ipm_regions = gpd.read_file('Data/IPM_Regions/national_emm_boundaries.shp')
+ipm_regions = gpd.read_file('AirPollution_Data/IPM_Regions_201770405.shp')
+zone_to_region = {
+    zone: region for region, zones in region_aggregations.items() for zone in zones
+}
+ipm_regions["model_region"] = ipm_regions["IPM_Region"].map(zone_to_region)
[…]

### 2025-05-23 Endogenous/Gen_Emissions_Locations_SimpleEmisCoefs.py

@@ -1,4 +1,5 @@
 # # Code to impute emissions rates, then output a file that specifies emissions per location per MWh of generation
+
 # The first challenge will be to get emissions rates (lb/mwh).  To do this I'll be taking EIA data on net generation and EPA data on emissions to calculate rates.  There will be many missings that I will fill in by taking capacity-weighted technology-planning_year means.
 # 
 # Once I have emissions rates, the goal will be to output a series of shapefiles of location-specific emissions per MWh. A challenge here is assigning output to to specific plants.  I assign production based on capacity.
@@ -28,24 +29,25 @@ import fiona
 import math
 import zipfile
 from shapely.geometry import Point
+import yaml
 
 #os.chdir('C:/Users/lbeatty/Documents/Lauren_MIP_Contribution/')
 os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
+
 # READ EVERYTHING IN
+from MIP_AirPollution.Endogenous.settings import *
+
+with open("MIP_results_comparison/case_settings/26-zone/settings-atb2023/model_definition.yml", "r") as file:
+    region_aggregations = yaml.safe_load(file)["region_aggregations"]
+region_aggregations['FRCC']=['FRCC']
 
 existing_gen_units = pd.DataFrame()
 scenario='current_policies_short'
 
-year_inputs = {
-    'p1': 2027,
-    'p2': 2030
-}
-
-
 #check if folder exists, and if not, unzip corresponding folder
-if not os.path.exists('Data/GenX inputs/'+scenario+'/'):
-    with zipfile.ZipFile('Data/GenX inputs/'+scenario+'.zip', 'r') as zip_ref:
-        zip_ref.extractall('Data/GenX inputs/'+scenario+'/')
+if not os.path.exists('AirPollution_Data/GenX inputs/'+scenario+'/'):
+    with zipfile.ZipFile('AirPollution_Data/GenX inputs/'+scenario+'.zip', 'r') as zip_ref:
+        zip_ref.extractall('AirPollution_Data/GenX inputs/'+scenario+'/')
 else:
 
     print('Data/GenX inputs/'+scenario+'/'+" already exists")
@@ -54,7 +56,7 @@ else:
[…]

### 2025-05-23 Endogenous/InMap_SR_GenObjCoefs.py

@@ -19,6 +19,7 @@ import os
 ## Code lifted (almost) directly from https://www.inmap.run/blog/2022/12/15/tutorial/
 
 os.chdir('C:/Users/laure/Documents/Switch-USA-PG/')
+from MIP_AirPollution.Endogenous.settings import *
 
 def rect(i, w, s, e, n):
     x = [w[i], e[i], e[i], w[i], w[i]]
@@ -45,14 +46,16 @@ sr = zarr.open(s3fs.S3Map(url, s3=fs, check=False), mode="r")
 
 p = poly(sr)
 
-years = ['2027', '2030', '2035', '2040']
-scenario = 'current_policies_short_simplified'
-simplified=True
+scenario = 'current_policies_short'
 
-for year in years:
-    filename='MIP_Air_OutData/MIP_Emissions/'+ scenario+ '_marginal_emissions_'+str(year)+'.shp'
+for year in year_inputs.values():
+    filename='MIP_Air_OutData/'+ scenario+ '_simplified_marginal_emissions_'+str(year)+'.shp'
 
     emis = gpd.read_file(filename)
+    emis = emis.query('SOx>0|PM2_5>0|VOC>0|NOx>0|NH3>0').reset_index()
+
+    #switch will place 0s for missings, so might as well drop zeros and reduce run-time
+
     fact = 28766.639
 
     df = pd.DataFrame({'Location': range(52411)})
@@ -122,7 +125,8 @@ for year in years:
     exposure_data_collapsed = exposure_data.groupby(['Race', 'Cluster','Pollutant']).agg({'Exposure':'sum'}).reset_index()
     exposure_data_collapsed['year']=year
 
-    exposure_data_collapsed.to_csv('MIP_Air_OutData/Marginal_Coefficients/' + scenario+'_marginal_exposure_coefs_'+year+'.csv')
+    year = str(year)
+    exposure_data_collapsed.to_csv('MIP_Air_OutData/' + scenario+'_marginal_exposure_coefs_'+year+'.csv')
 
 
 

### 2025-05-23 Get_CAMD.py

@@ -1,55 +1,0 @@
-import requests
-import json
-import pandas as pd
-
-# Set API key and base URL
-# API Keys can be requested at https://www.epa.gov/power-sector/cam-api-portal#/api-key-signup
-f = open("MIP_AirPollution/API_Key.json")
-config = json.load(f) 
-API_Key = config['API_Key']
-
-data_directory = "C:/Users/lbeatty/Documents/Lauren_MIP_Contribution/Data/"
-base_url = "https://api.epa.gov/easey/"
-
-# Function to fetch data
-def fetch_data(endpoint, year):
-    data = []
-    page = 1
-    while True:
-        url = f"{base_url}{endpoint}?page={page}&perPage=100&year={year}"
-        headers = {
-            "x-api-key": API_Key,
-            "accept": "application/json"
-        }
-        response = requests.get(url, headers=headers)
-        if response.status_code == 200:
-            content = response.json()
-            data.extend(content)
-            total_count = int(response.headers.get('X-Total-Count', 0))
-            if total_count <= page * 100:
-                break
-            page += 1
-        else:
-            print(f"Failed to fetch data for year {year}, page {page}")
-            break
-    return data
-
-# Fetch facilities attributes data
-facilities_attributes_data = []
-for year in range(2015, 2022):
-    endpoint = f"facilities-mgmt/facilities/attributes"
-    facilities_attributes_data.extend(fetch_data(endpoint, year))
-
-# Convert to DataFrame and save as CSV
-facilities_attributes_df = pd.DataFrame(facilities_attributes_data)
-facilities_attributes_df.to_csv(data_directory+"CAMD/facilities_attributes.csv", index=False)
-
-# Fetch facilities emissions data
-facilities_emissions_data = []
-for year in range(2020, 2022):
-    endpoint = f"emissions-mgmt/emissions/apportioned/annual"
-    facilities_emissions_data.extend(fetch_data(endpoint, year))
-
-# Convert to DataFrame and save as CSV
-facilities_emissions_df = pd.DataFrame(facilities_emissions_data)
-facilities_emissions_df.to_csv(data_directory+"CAMD/facilities_emissions.csv", index=False)

### 2025-05-23 download_data.py

@@ -74,7 +74,10 @@ urls = [
     "https://www2.census.gov/programs-surveys/popest/tables/2020-2024/counties/totals/co-est2024-pop.xlsx",
     "https://www.eia.gov/electricity/data/eia860/archive/xls/eia8602022.zip",
     "https://www.eia.gov/electricity/data/eia923/archive/xls/f923_2022.zip",
-    "https://gaftp.epa.gov/Air/emismod/2020/2020emissions/2020ha2_point_inventory_22sep2023.zip"
+    "https://gaftp.epa.gov/Air/emismod/2020/2020emissions/2020ha2_point_inventory_22sep2023.zip",
+    "https://github.com/USEPA/camd-eia-crosswalk/releases/download/v0.3/epa_eia_crosswalk.csv",
+    'https://www.epa.gov/system/files/documents/2024-06/egrid-draft-pm-emissions.xlsx',
+    'https://www.epa.gov/sites/production/files/2019-08/ipm_v6_regions.zip'
 ]
 
 for url in urls:

### 2025-05-23 Paper_Endogenous.tex

@@
 \author{Lauren Beatty\thanks{Environmental Defense Fund  \hspace{.5cm} E-mail: lbeatty1@edf.org \hspace{.5cm}Website: \href{https://lbeatty1.github.io}{https://lbeatty1.github.io}}\\
-Matthias Fripp\thanks{Environmental Defense Fund}}
+Matthias Fripp\thanks{Energy Innovations}}
 
@@
 \section{Introduction}
-Climate change mitigation will require deep cuts to greenhouse gas emissions from the electricity sector.
+Climate change mitigation requires deep cuts to greenhouse gas emissions from the electricity sector.
 Least-cost electricity planning models are powerful tools that help policymakers and capacity planners find portfolios of generation technologies that meet reliability requirements and emissions goals.
@@
 One major external costs is the generation of local air pollutants.
-There is a large body of work on the health effects of air pollution, and a large body of work on the effects of electricity generation on air pollution.
+There is a large body of work on the health effects of air pollution, as well as a large body of work on the effects of electricity generation on air pollution.
 Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acute lower respiratory illness, cerebrovascular disease, ischaemic heart disease, and lung cancer.
@@
 \cite{Thind2019FineGeography} find that Black and non-Hispanic white people are exposed to more pollution than other groups and that exposures for low-income people are higher than for high-income people, even after controlling for race.
-This highlights the importance for developing planning methods that alleviate rather than reinforce disproportionate burdens.
+This highlights the importance of developing planning methods that alleviate rather than reinforce disproportionate burdens.
 
@@
 
-Least-cost electricity planning models are the gold-standard models used to model capacity expansion in policy.
+Least-cost electricity planning models are the gold standard models used to model capacity expansion in policy.
 However, these models tend to have a narrow focus on minimizing costs -- potentially at the expense of distributional concerns.
@@
 
-This paper makes an important contribution to the environmental justice literature by ...
+This paper makes an important contribution to the environmental justice literature by endogenizing air pollution within a least-cost planning model, and proposing a framework to equitably reduce air pollution.
+Our findings indicate that major improvements to air pollution can be made with relatively small increases in overall cost.
 
 \section{Data}
+The fist task in our method is to construct emissions rates for sources in the model in tons per MWh.
+These emissions rates are constructed by combining emissions data from the U.S.
+Environmental Protection Agency (EPA), and generations data from the U.S.
+Energy Information Administration (EIA).
+We use the EPA-EIA Power Sector Data Crosswalk to merge generators across the two sources \citep{HuettemanEPA-EIACrosswalk}.
+
 \subsection{Emissions Data}
@@
 I get emissions of NH$_3$, VOC, and PM2.5 from the NEI.
+Air pollution data is used to calculate source emissions rates in tons per MWh.
+
+\subsection{Generation Data}
+Data on the generation and location of existing plants comes from the EIA.
 
@@
 \section{Results}
+\subsection{Model Inputs}
+There are a few findings from the construction of model inputs that are worth discussing.
+First, unsurprisingly, empirical data on emissions and generation reveals that coal is a dirty source of generation.
+However, when we translate emissions into marginal exposure coefficients, we find that the average relative difference between coal and natural gas decreases.
+In particular, it appears that natural gas plants are on average, closer to Asian and Latino population centers.
+A unit of electricity generated from coal only causes slightly more damage (in terms of population-weighted PM2.5 exposure) to Asians and Latinos than a unit of electricity generated from natural gas.
+In figure \ref{EmissionsExposureCoefs} we plot the average empirical emissions rates by technology in panel (a), and the mean coefficients describing how one MWh of generation affects mean group-level exposure of PM2.5 in $\mu$g/m$^3$.
+
+This start divergence between the relative emissions rates and the relative exposure coefficients highlights one of the central premises of the paper -- that where generation is placed can have large impacts on pollution exposure, and that selectively choosing what generation goes where can lead to large decreases in pollution exposure at low costs.
 
@@
 \end{figure}
-In figure \ref{EmissionsExposureCoefs} we plot both average emissions rates, and average exposure coefficients for various technologies.
+
+\subsection{Model Results}
+Our high-level, key finding is that even a modest budget of $\$50$ billion through 2050 could pay for massive reductions in equitable air quality improvement.
+Our results have features that you would expect based on economic theory.
+There are increasing costs of reducing air pollution -- the gains from moving from least-cost to allocating $\$50$ billion are much larger than the gains from moving from allocating $\$50$ billion to $\$100$ billion.
+Our results also suggest that at the optimum, the min-max objective generates equal average exposures for all groups except Native Americans, who have consistently lower exposure levels.
 
@@
 
+\subsection{Comparison with carbon-only cuts}
+While we think these results have important policy implications, namely that improving air quality equitably can happen at low costs, it's important to note that our results don't point to which policies would achieve the optimal solution.
+Rather, these results are prescriptive -- \textit{"Build X, Y, and Z in regions A,B, and C to achieve the deepest equitable cuts in air pollution at a certain cost."}
+
+Previous papers have found that carbon policies tend to decrease both total air pollution exposure, and the inequalities in air pollution exposure.
+This begs the question, how does a portfolio that minimizes air pollution exposure compare to a least-cost portfolio aimed at reducing carbon?
+
 \section{Conclusion}

### 2025-05-23 Slides_3_6.tex

--- 2024-10-18 Slides_5_3_2024.tex
@@
 %Information to be included in the title page:
-\title{Designing Equitable Electricity Systems}
-\author{Lauren Beatty\\ \href{mailto:lbeatty@edf.org}{lbeatty@edf.org}\\
-\vspace{.5cm}\\
-Matthias Fripp\\
-\href{mailto:mfripp@edf.org}{mfripp@edf.org}}
+\title{Air Pollution from Electricity Production}
+\author{Lauren Beatty\\ \href{mailto:lbeatty@edf.org}{lbeatty@edf.org}}
 \institute{Environmental Defense Fund}
-\date{May 6, 2024}
+\date{\today}
 
@@
 
-\logo{\includegraphics[height=1cm]{EDF_color_webRGB_2022 (2).png}\hspace*{.03\paperwidth}\vspace{.01\paperwidth}}
+\logo{\includegraphics[height=.5cm]{EDF_color_webRGB_2022 (2).png}\hspace*{.03\paperwidth}\vspace{.01\paperwidth}}
 
@@
 \frametitle{Introduction}
-\textbf{Main Question:  How do we design electrical systems that center equity and justice \textit{without} sacrificing efficiency?}
+Two Different Questions / Approaches:
+\begin{enumerate}
+    \item Downscaling
+    \begin{itemize}
+        \item \textbf{Question}: How might a particular plan for decarbonization affect air pollution.
+        \item \textbf{Method}: Sequentially run a least-cost planning model, then take the outputs of that model and predict pollution.
+    \end{itemize}
+    \item Endogenous
+    \begin{itemize}
+        \item \textbf{Question}: What decarbonization plan most dramatically decreases air pollution equitably?
+        \item \textbf{Method}: Alter the planning model to minimize pollution subject to a cost constraint.
+    \end{itemize}
+\end{enumerate}
 \vspace{1cm}
 
-\paragraph{\textbf{Previous Paradigm}}
-\begin{itemize}
-    \item Potentially harmful and polluting infrastructure is sited disproportionately near poor and marginalized communities.
-    \item Planning decisions are informed by \textit{least cost} models.
-    \item Communities are left in the lurch by technological changes.
-\end{itemize}
 \end{frame}
@@
 
-{\nologo
-\begin{frame}{Previous Attempts to Shoehorn Justice Into these Models}
-\begin{center}
-\begin{tikzpicture}[node distance=2cm]
+\begin{frame}{Background on Least-Cost Electricity Planning Models: Pros and Cons}
+    \begin{itemize}
+        \item Powerful for out-of-sample questions
 
-\node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
-• Constraints}};
+        \item Persuasive in policy arguments.
 
-\node(obj)[process, below of=start, text width=8cm]{\shortstack[l]{$\min \sum \text{Costs} + \text{Monetized Air Pollution Impacts}$ \\ \textit{ 
-   s.t.   }\text{Everyone gets electricity and nothing breaks }}};
+        \item No or very little backwards validation -- We have no evidence that these models make good predictions about what will happen.
 
-\node (result) [stop, below of=obj] {\shortstack[l]{Outputs\\• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};
+        \item Rely on strong assumptions about future costs and demand, as well as the strong assumption that these are known ex-ante.
+    \end{itemize}
+\end{frame}
 
-\draw [arrow] (start) -- (obj);
-\draw [arrow] (obj) -- (result);
+\begin{frame}{Air Pollution Modelling}
+\textbf{Translate emissions at points to concentrations over space.}
 
-\end{tikzpicture}
-\end{center}
+\vspace{.3cm}
 
+Multiple models of varying complexity. The main models used in policy work are:
 \begin{itemize}
-    \item Monetized impacts don't consider unjust distributional consequences.
-    \item Equity weights can be difficult to generate and defend.
+    \item APEEP
+    \item INMAP
+    \item EASIUR
 \end{itemize}
 
-\end{frame}}
+\vspace{.3cm}
+
+These are all \textbf{average-annual models}, with reduced-form representations of more complex chemical transport models.
+Each answers the question: \textit{How much does pollution affect average annual exposure?} Therefore, they are suitable for thinking about policies affecting aggregate emissions, but not suitable for thinking about things like hourly NAAQs.
+\end{frame}
+
+\begin{frame}{Downscaling}
+\begin{enumerate}
+    \item Run least-cost planning model
+    \item Model outputs annual generation for each ``resource," such as Combined Cycle Plants in New England ISO North.
+    \begin{itemize}
+        \item How much resources are aggregated is a user choice which needs to balance computational feasibility and model accuracy.
+    \end{itemize}
+    \item Translate model outputs into point source locations and aggregate emissions of important pollutants (in InMAP -- PM2.5, NOx, SOx, VOCs, NH3).
+        \begin{itemize}
+            \item Involves lots of assumptions about how to allocate generation to existing plants, and the location of new plants.
+        \end{itemize}
+    \item Run those aggregate emissions through an air pollution model of your choice.
+    \item Translate concentrations into health impacts.
+\end{enumerate}
+\end{frame}
+
+\begin{frame}{MIP Results}
+    \begin{columns} % Create two columns
+        \column{0.5\textwidth} % First column: 50% of slide width
+        \centering
+        \includegraphics[width=\linewidth]{Figures/Output/full-base-200/GenX/ISRM_2027_TotalPM25_concentrationmap.jpg}
+        \captionof{figure}{PM2.5 Concentrations in 2027 in the base-200 scenario}
+
+        \column{0.5\textwidth} % Second column: 50% of slide width
+        \centering
+        \includegraphics[width=\linewidth]{Figures/Output/full-base-200/GenX/ISRM_2050_TotalPM25_concentrationmap.jpg}
+        \captionof{figure}{PM2.5 Concentrations in 2050 in the base-200 scenario}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{MIP Results}
+    \includegraphics[width=\textwidth]{Figures/Output/Compare_scenarios_current-base_GenX.jpg}
+\end{frame}
 
 {\nologo
-\begin{frame}{Our Approach}
+\begin{frame}{Endogenous Exposure}
     \begin{center}
@@
         \item Use InMap Source-Receptor Matrix to calculate how one MWh of generation for each resource affects population-weighted average exposure for different groups.
-        \item Use Switch to calculate least cost planning costs.
+        \item Use Switch to calculate least-cost planning costs.
         \item Modify and run Switch to minimize air pollution exposure subject to a cost constraint.
@@
     \vspace{.5cm}
-    $\Rightarrow$ Perhaps you can get large improvements in health or jobs at only modest increases in cost.
+    $\Rightarrow$ Perhaps you can get large improvements in health at only modest increases in cost.
+\end{frame}
+
+\begin{frame}{Flavor of Results}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Figures/EndogenousResults/current_policies_short_simplified/exposure_cost_PPF.png}
 \end{frame}
@@
 \end{frame}
-
-\begin{frame}{Bonus Slide!}
-    This project started as a side-project to the Model Intercomparison Project -- a project aimed at harmonizing the inputs and comparing multiple open-source electricity planning models.
-
-    \vspace{.4cm}
-
-    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2030.jpg}
-    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2040.jpg}
-    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2050.jpg}
-
-\begin{itemize}
-    \item Death rates vary widely across racial groups.
-    \item Aggressive carbon policies reduce pollution exposure the fastest.
-    \item Optimistic -- even under current policies future generations should have much better air quality.
-\end{itemize}
-
-\end{frame}
-
 \end{document}

### 2025-06-02 Endogenous/Paper_figures.py
working on scales for figs

@@ -85,6 +85,12 @@ def mix(color1, color2, ratio):
     return mcolors.to_hex((1 - ratio) * c1 + ratio * c2)
 
 light_red = mix(dark_red, white, 0.5)
+lighter_blue = mix(dark_blue, white, 0.3)
+
+def round_up_to_step(high, low, step):
+    steps = math.ceil((high - low) / step)
+    return low + steps * step
+
 
 #read in exposure coefs
 coefs=pd.DataFrame()
@@ -458,6 +464,7 @@ fig.savefig("MIP_AirPollution/Figures/EndogenousPaper/wind_capacity_map.png", dp
 ###################
 ## Plot Dispatch ##
 ###################
+
 dispatch_spatial = dispatch.groupby(['gen_energy_source', 'gen_load_zone', 'period', 'scenario']).agg({'Energy_GWh_typical_yr':'sum'}).reset_index()
 
 base = dispatch_spatial[dispatch_spatial['scenario']==f'{scenario}']
@@ -466,30 +473,42 @@ base.rename(columns={'Energy_GWh_typical_yr': 'Base_Gen'}, inplace=True)
 
 dispatch_spatial = pd.merge(dispatch_spatial, base, on=['gen_load_zone', 'gen_energy_source', 'period'], how='left')
 
-#calculate difference in capacity
+#calculate difference in dispatch
 dispatch_spatial['dif']=dispatch_spatial['Energy_GWh_typical_yr']-dispatch_spatial['Base_Gen']
-dispatch_spatial = dispatch_spatial.merge(ipm_regions, right_on='model_region', left_on='gen_load_zone', how='left')
+
+## first, figure out what big, small differences in generation are
+
+max_dif = dispatch_spatial['dif'].max()
+min_dif = dispatch_spatial['dif'].min()
+max_dif = max(abs(max_dif), abs(min_dif))
 
 c=200
 y=2050
 tech = 'Solar'
 
-plotdata = capacity_spatial[capacity_spatial['scenario']==f'{scenario}_{c}']
-plotdata = plotdata[plotdata['PERIOD']==y]
+plotdata = dispatch_spatial[dispatch_spatial['scenario']==f'{scenario}_{c}']
+plotdata = plotdata[plotdata['period']==y]
 plotdata = plotdata[plotdata['gen_energy_source']==tech]
+
+plotdata = ipm_regions.merge(plotdata, left_on='model_region', right_on='gen_load_zone', how='left')
+
[…]

### 2025-06-03 Endogenous/Paper_figures.py
a few more plots

@@ -6,11 +6,16 @@ import os
 import numpy as np
 import pandas as pd
 import math
+import seaborn as sns
 from matplotlib.colors import LinearSegmentedColormap
 from matplotlib.colors import ListedColormap
 import matplotlib.pyplot as plt
 from matplotlib.colors import LinearSegmentedColormap, Normalize
 import yaml
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+from shapely.geometry import LineString
+# import matplotlib.colors as mcolors
+
 
 
 
@@ -138,6 +143,14 @@ for c in cost_limits:
     tempdat = pd.read_csv(f"{result_dir}{scenario}_{c}/GroupExposure.csv")
     tempdat['scenario']=f'{scenario}_{c}'
     group_exposure = pd.concat([group_exposure,tempdat])
+
+## transmission
+transmission = pd.read_csv(f"{result_dir}{scenario}/transmission.csv")
+transmission['scenario']=scenario
+for c in cost_limits:
+    tempdat = pd.read_csv(f"{result_dir}{scenario}_{c}/transmission.csv")
+    tempdat['scenario']=f'{scenario}_{c}'
+    transmission = pd.concat([transmission, tempdat])
 
 ## Plot coefs and emissions
 coefs['technology'] = coefs.apply(update_technology, axis=1, column='Cluster')
@@ -765,3 +778,161 @@ cbar.set_label("Difference (GWh)")
 ax.set_title("Coal Generation in 2027 with $200B Budget \n Relative to Base", fontsize=15)
 ax.set_axis_off()
 fig.savefig("MIP_AirPollution/Figures/EndogenousPaper/coal_generation_map_2027_50B.png", dpi=300, bbox_inches='tight')
+
+##################
+## Transmission ##
+##################
+base = transmission[transmission['scenario']==f'{scenario}']
+base = pd.DataFrame(base[['PERIOD', 'trans_lz1', 'trans_lz2', 'TxCapacityNameplateAvailable']])
+base.rename(columns={'TxCapacityNameplateAvailable': 'Base_Capacity'}, inplace=True)
+
+transmission = pd.merge(transmission, base, how='left', on=['PERIOD', 'trans_lz1', 'trans_lz2'])
+c=200
+y=2050
+plotdata = transmission[transmission['PERIOD']==y]
+plotdata = plotdata[plotdata['scenario']==f'{scenario}_{c}']
[…]

### 2025-06-03 Paper_Endogenous.tex

@@
     \centering
-    \subcaptionbox{2027}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/Dispatch_Relative_by_scenario_2027.png}}
-    \subcaptionbox{2030}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/Dispatch_Relative_by_scenario_2030.png}}
+    \subcaptionbox{2027}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/current_policies_short/Dispatch_Relative_by_scenario_2027.png}}
+    \subcaptionbox{2030}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/current_policies_short/Dispatch_Relative_by_scenario_2030.png}}\\
+    \subcaptionbox{2035}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/current_policies_short/Dispatch_Relative_by_scenario_2035.png}}
+    \subcaptionbox{2040}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/current_policies_short/Dispatch_Relative_by_scenario_2040.png}}\\
+    \subcaptionbox{2045}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/current_policies_short/Dispatch_Relative_by_scenario_2045.png}}
+    \subcaptionbox{2050}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/current_policies_short/Dispatch_Relative_by_scenario_2050.png}}\\
     \caption{Generation by Technology by Scenario}
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{Figures/EndogenousResults/base_short/exposure_cost_PPF.png}
+    \includegraphics[width=0.7\linewidth]{Figures/EndogenousResults/current_policies_short/exposure_cost_PPF.png}
     \caption{Here we show the relationship between the objective function (min-max of exposure) and the cost limit constraint.}

### 2025-06-03 Endogenous/Paper_figures.py
adjusting look of plots

@@ -67,11 +67,11 @@ zone_to_region = {
 
 scenario_mapping = {
     'current_policies_short': 'Base Case',
-    'current_policies_short_50': '$50 Billion Budget',
-    'current_policies_short_100': '$100 Billion Budget',
-    'current_policies_short_200': '$200 Billion Budget'
+    'current_policies_short_50': '$50 Billion',
+    'current_policies_short_100': '$100 Billion',
+    'current_policies_short_200': '$200 Billion'
 }
-scenario_order = ['Base Case', '$50 Billion Budget', '$100 Billion Budget', '$200 Billion Budget']
+scenario_order = ['Base Case', '$50 Billion', '$100 Billion', '$200 Billion']
 
 # Get the RdBu colormap
 rd_bu = plt.cm.get_cmap('RdBu')
@@ -176,17 +176,18 @@ bar4 = ax.bar(index + 3*bar_width, emis_summary['NH3'], bar_width, label='NH3', 
 bar5 = ax.bar(index + 4*bar_width, emis_summary['VOC'], bar_width, label='VOC', color=tab10[4])
 
 # Labels and title
-ax.set_xlabel('Technology', fontsize=14, fontweight='bold')
-ax.set_ylabel('Mean Emissions (Tons/MWh)', fontsize=14, fontweight='bold')
-ax.set_title('Mean Emissions by Technology', fontsize=16, fontweight='bold')
+ax.set_xlabel('Technology', fontsize=16, fontweight='bold')
+ax.set_ylabel('Mean Emissions (Tons/MWh)', fontsize=18, fontweight='bold')
+ax.set_title('Mean Emissions by Technology', fontsize=18, fontweight='bold')
 
 
 # Rotate x-axis labels
 ax.set_xticks(index + 2*bar_width)
 ax.set_xticklabels(emis_summary['technology'], rotation=0, ha='center')
+ax.tick_params(axis='x', labelsize=16)
 
 # Legend and layout
-ax.legend()
+ax.legend(fontsize=16)
 plt.tight_layout()
 
 # Save the figure
@@ -213,16 +214,17 @@ ax.bar(index + 3*bar_width, exposure_wide['Native'], bar_width, label='Native Am
 ax.bar(index + 4*bar_width, exposure_wide['WhiteNoLat'], bar_width, label='White', color=tab10[4])
 
 # Labels and title (bold + larger font)
-ax.set_xlabel('Technology', fontsize=14, fontweight='bold')
[…]

### 2025-06-03 weai.tex

--- 2024-10-18 Slides_5_3_2024.tex
@@
 \usepackage{hyperref}
+\usepackage{natbib}
 \usepackage{color}
+\usepackage{changepage} % in your preamble
+
 %Information to be included in the title page:
-\title{Designing Equitable Electricity Systems}
-\author{Lauren Beatty\\ \href{mailto:lbeatty@edf.org}{lbeatty@edf.org}\\
-\vspace{.5cm}\\
+\title{AERE @ WEAI\\
+Rawls for Environmental Justice}
+\author{Lauren Beatty\\ \href{mailto:lbeatty@edf.org}{lbeatty@edf.org} \\
+\vspace{.3cm}
 Matthias Fripp\\
-\href{mailto:mfripp@edf.org}{mfripp@edf.org}}
+\href{mailto:matthias@energyinnovation.org}{matthias@energyinnovation.org}}
 \institute{Environmental Defense Fund}
-\date{May 6, 2024}
+\date{June 21, 2024}
 
@@
 
-\begin{frame}
-\frametitle{Introduction}
-\textbf{Main Question:  How do we design electrical systems that center equity and justice \textit{without} sacrificing efficiency?}
-\vspace{1cm}
+\section{Introduction}
 
-\paragraph{\textbf{Previous Paradigm}}
+\begin{frame}{Introduction}
 \begin{itemize}
-    \item Potentially harmful and polluting infrastructure is sited disproportionately near poor and marginalized communities.
-    \item Planning decisions are informed by \textit{least cost} models.
-    \item Communities are left in the lurch by technological changes.
+    \item Air pollution is responsible for $\sim 7$ million deaths annually (\href{https://www.who.int/health-topics/air-pollution#tab=tab_2}{WHO}).
+    \item In the U.S.,  Black and non-Hispanic white people are exposed to higher levels of air pollution than other groups.
+Low income people are exposed to higher pollution levels than high income people, controlling for race \citep{Thind2019FineGeography}.
+    \item \citet{Lelieveld2015TheScale} estimate that power generation is linked with 31$\%$ of premature mortality caused by outdoor air pollution in the U.S.
+\end{itemize}
+
+    \textbf{How can we reduce air quality \textit{equitably?}}
+\end{frame}
+
+\begin{frame}{Previous Work}
+\begin{itemize}
+    \item \citet{Sergi2020OptimizingBenefits} include co-benefits involved with reducing electricity generation from fossil fuels in the objective function.
+    \item citep{Goforth2022AirStrategies} and \citep{Shawhan2024PoliciesAmericans} project how the electricity transmission might affect the distributional consequences of air pollution exposure.
 \end{itemize}
@@
 
-\begin{frame}{Background on Least-Cost Electricity Planning Models}
+\begin{frame}{Our Approach}
+    \begin{enumerate}
+        \item Use EPA emissions data to calculate emissions rates (Pounds per MWh)
+        \item Use InMap Source-Receptor Matrix to calculate how one MWh of generation for each resource affects population-weighted average exposure for different groups.
+        \item Use Switch to calculate least-cost planning costs.
+        \item Modify and run Switch to minimize air pollution exposure subject to a cost constraint.
+    \end{enumerate}
+
+\paragraph{Benefits}
+\begin{itemize}
+    \item Avoids the use of equity weights while still centering distributional concerns.
+    \item Still finds low-cost solutions.
+    \item Generalizable framework
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Least-Cost Electricity Planning Models}
 \begin{center}
@@
 {\nologo
-\begin{frame}{Previous Attempts to Shoehorn Justice Into these Models}
-\begin{center}
-\begin{tikzpicture}[node distance=2cm]
-
-\node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
-• Constraints}};
-
-\node(obj)[process, below of=start, text width=8cm]{\shortstack[l]{$\min \sum \text{Costs} + \text{Monetized Air Pollution Impacts}$ \\ \textit{ 
-   s.t.   }\text{Everyone gets electricity and nothing breaks }}};
-
-\node (result) [stop, below of=obj] {\shortstack[l]{Outputs\\• Time-varying generation for each resource \\ • New Generation Capacity \\ • New Transmission Capacity}};
-
-\draw [arrow] (start) -- (obj);
-\draw [arrow] (obj) -- (result);
-
-\end{tikzpicture}
-\end{center}
-
-\begin{itemize}
-    \item Monetized impacts don't consider unjust distributional consequences.
-    \item Equity weights can be difficult to generate and defend.
-\end{itemize}
-
-\end{frame}}
-
-{\nologo
-\begin{frame}{Our Approach}
+\begin{frame}{Endogenous Exposure}
     \begin{center}
@@
 
-\begin{frame}{Our Approach}
-    \begin{enumerate}
-        \item Use EPA emissions data to calculate emissions rates (Pounds per MWh)
-        \item Use InMap Source-Receptor Matrix to calculate how one MWh of generation for each resource affects population-weighted average exposure for different groups.
-        \item Use Switch to calculate least cost planning costs.
-        \item Modify and run Switch to minimize air pollution exposure subject to a cost constraint.
-    \end{enumerate}
-
-\paragraph{Benefits}
-\begin{itemize}
-    \item Avoids the use of equity weights while still centering distributional concerns.
-    \item Still finds low-cost solutions.
-    \item Generalizable framework
-\end{itemize}
-\end{frame}
-
 \begin{frame}{A Point About Calculus}
@@
     \vspace{.5cm}
-    $\Rightarrow$ Perhaps you can get large improvements in health or jobs at only modest increases in cost.
+    $\Rightarrow$ Perhaps you can get large improvements in health at only modest increases in cost.
+\end{frame}
+
+\begin{frame}{Data}
+
+\end{frame}
+
+\begin{frame}{Results}
+    \begin{itemize}
+        \item Imperfect correlation between emissions rates, and effects on pollution.
+        \item Modest budgets can dramatically reduce air pollution exposure from electricity generation.
+        \item Air quality improvements exhibit decreasing marginal returns.
+        \item The mechanisms for reducing air pollution are:
+        \begin{itemize}
+            \item Expanding wind and solar
+            \item Decreasing coal generation faster
+            \item Changing the spatial distribution of remaining fossil generation
+            \item Increases in transmission
+        \end{itemize}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Imperfect Correlation Between Emissions Rates and Effects on Pollution}
+\includegraphics[width=0.49\textwidth]{Figures/EndogenousPaper/emissions_by_technology.png}
+\includegraphics[width=0.49\textwidth]{Figures/EndogenousPaper/exposure_by_technology.png}
+\end{frame}
+
+\begin{frame}{Large Improvements in Air Quality at Modest Costs}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Figures/EndogenousPaper/exposure_by_scenario_2050.png}
+\end{frame}
+
+\begin{frame}{Decreasing Marginal Returns}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Figures/EndogenousPaper/exposure_cost_PPF.png}
+\end{frame}
+
+\begin{frame}{Decreasing Fossil and Increasing Renewable Generation}
+\begin{adjustwidth}{-3em}{-1.5em}
+
+    \centering
+    \begin{minipage}{1.1\textwidth}
+        \centering
+        \includegraphics[width=0.32\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2027.png} \hfill
+        \includegraphics[width=0.32\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2030.png} \hfill
+        \includegraphics[width=0.32\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2035.png}
+    \end{minipage}
+
+    \vspace{0.5cm}
+
+    \begin{minipage}{1.1\textwidth}
+        \centering
+        \includegraphics[width=0.32\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2040.png} \hfill
+        \includegraphics[width=0.32\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2045.png} \hfill
+        \includegraphics[width=0.32\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2050.png}
+    \end{minipage}
+\end{adjustwidth}
+\end{frame}
+
+\begin{frame}{Increasing Renewable Generation}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Figures/EndogenousPaper/solar_generation_map.png}
+\end{frame}
+
+\begin{frame}{Increasing Renewable Generation}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Figures/EndogenousPaper/wind_generation_map.png}
+\end{frame}
+
+\begin{frame}{Decreasing Coal Faster}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Figures/EndogenousPaper/coal_generation_map_2027_200B.png}
+\end{frame}
+
+\begin{frame}{Changing the Spatial Distribution of Remaining Fossil Generation}
+\centering
+\includegraphics[width=0.8\textwidth]{Figures/EndogenousPaper/naturalgas_generation_map.png}
+\end{frame}
+\begin{frame}{Transmission/Capacity Substitution}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/EndogenousPaper/transmission_capacity_substitution.png}
 \end{frame}
@@
 
-\begin{frame}{Bonus Slide!}
-    This project started as a side-project to the Model Intercomparison Project -- a project aimed at harmonizing the inputs and comparing multiple open-source electricity planning models.
-
-    \vspace{.4cm}
-
-    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2030.jpg}
-    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2040.jpg}
-    \includegraphics[width=0.3\textwidth]{Figures/Output/ISRM_deathrate_by_scenario_2050.jpg}
-
-\begin{itemize}
-    \item Death rates vary widely across racial groups.
-    \item Aggressive carbon policies reduce pollution exposure the fastest.
-    \item Optimistic -- even under current policies future generations should have much better air quality.
-\end{itemize}
-
+\begin{frame}{References}
+\newpage
+\bibliographystyle{jpe}
+%\bibliographystyle{econometrica}
+\bibliography{references.bib}
 \end{frame}
-
 \end{document}

### 2025-06-03 Endogenous/Paper_figures.py
update legend

@@ -265,16 +265,31 @@ for year in sorted(years, reverse=True):
     if year == 2050:
         # Get handles and labels from the original axis
         handles, labels = ax.get_legend_handles_labels()
-        
+
         # Create a new figure for the legend
         fig_legend = plt.figure(figsize=(8, 6))
-        fig_legend.legend(handles, labels, loc='center', fontsize=18, frameon=False, ncol=1)  # Customize as needed
-        fig_legend.tight_layout()
         
-        # Save legend separately
-        fig_legend.savefig(f'MIP_AirPollution/Figures/EndogenousPaper/Legend_{year}.png', dpi=300, bbox_inches='tight')
-        plt.close(fig_legend)  # Close legend figure
+        # Add legend and store it in a variable
+        legend = fig_legend.legend(
+            handles, labels,
+            loc='center',
+            fontsize=18,
+            frameon=False,
+            ncol=2,
+            title='Technology'
+        )
+        
+        # Set legend title font size
+        legend.get_title().set_fontsize(24)
 
+        # Save the figure with the legend
+        fig_legend.savefig(
+            f'MIP_AirPollution/Figures/EndogenousPaper/Legend_{year}.png',
+            dpi=300,
+            bbox_inches='tight'
+        )
+        plt.close(fig_legend)
+        
     ax.get_legend().remove()
     plt.tight_layout()
     plt.savefig(f'MIP_AirPollution/Figures/EndogenousPaper/Dispatch_Relative_by_scenario_{year}.png', dpi=300, bbox_inches='tight')  # DPI for high resolution

### 2025-06-06 Endogenous/Paper_figures.py
fixed emissionsexposure plot

@@ -966,5 +966,5 @@ plt.title('CO2 Emissions versus Max(Group Exposure)')
 plt.grid(True)
 plt.tight_layout()
 plt.legend(bbox_to_anchor=(1.05, 1), loc='upper left')
+plt.savefig("MIP_AirPollution/Figures/EndogenousPaper/emissions_vs_exposure.png", dpi=300, bbox_inches='tight')
 plt.show()
-fig.savefig("MIP_AirPollution/Figures/EndogenousPaper/emissions_vs_exposure.png", dpi=300, bbox_inches='tight')

### 2025-06-06 Paper_Endogenous.tex

--- 2024-12-02 Paper_Endogenous.tex
@@
 This method provides a menu of plans that equitably reduce pollution exposure.
+Our model results show that a modest $\$$50 billion investment through 2050 could approximately cut mean air pollution exposures in half and generate the largest reductions for groups with the highest baseline exposures.
 \end{abstract}
@@
 \section{Introduction}
-Climate change mitigation will require deep cuts to greenhouse gas emissions from the electricity sector.
-Least-cost electricity planning models are powerful tools that help policymakers and capacity planners find portfolios of generation technologies that meet reliability requirements and emissions goals.
+Mitigating climate change requires deep cuts in greenhouse gas emissions from the electricity sector.
+Least-cost electricity planning models have become essential tools for policymakers and utility planners seeking cost-effective generation portfolios that meet reliability requirements and emissions goals.
 
-However, these models have historically ignored some of the external costs of electricity generation.  
- One major external costs is the generation of local air pollutants.
-There is a large body of work on the health effects of air pollution, and a large body of work on the effects of electricity generation on air pollution.
-Air pollution is linked with numerous health impacts such as chronic obstructive pulmonary disease (COPD), acute lower respiratory illness, cerebrovascular disease, ischaemic heart disease, and lung cancer.
-\cite{Lelieveld2015TheScale} estimate that power generation is linked with 31$\%$ of premature mortality caused by outdoor air pollution in the U.S.
+However, these models often ignore external costs associated with electricity generation—especially the health damages from local air pollution.
+Air pollution from fossil fuel combustion has been linked to a wide range of adverse health outcomes, including chronic obstructive pulmonary disease (COPD), cerebrovascular disease, ischemic heart disease, and lung cancer.
+In the U.S., \citet{Lelieveld2015TheScale} estimate that power generation contributes to 31$\%$ of premature mortality caused by outdoor air pollution.
 
-There is also work that has shown that air pollution exposures vary drastically across racial and socioeconomic groups.
-\cite{Thind2019FineGeography} find that Black and non-Hispanic white people are exposed to more pollution than other groups and that exposures for low-income people are higher than for high-income people, even after controlling for race.
-This highlights the importance for developing planning methods that alleviate rather than reinforce disproportionate burdens.
+These harms are not evenly distributed.
+Numerous studies, including \citet{Thind2019FineGeography}, show that pollution exposure varies substantially across racial and socioeconomic groups.
+In the U.S., Black and white populations are exposed to higher pollution levels than other groups, and low-income populations face disproportionately high exposure, even after controlling for race.
+These disparities raise important equity concerns as the energy system transitions away from fossil fuels.
 
-Previous work has included co-benefits involved with reducing electricity generation from fossil fuels in the objective function \citet{Sergi2020OptimizingBenefits}.
-or projected how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}.
+Some prior studies have quantified the health co-benefits of decarbonization by monetizing pollution damages and incorporating them into the objective function of least-cost planning models \citep{Sergi2020OptimizingBenefits}).
+Others have explored how the electricity transition will affect the spatial distribution of air pollution and its consequences (\citet{Shawhan2024PoliciesAmericans}; \citet{Goforth2022AirStrategies}).
 
-Least-cost electricity planning models are the gold-standard models used to model capacity expansion in policy.
-However, these models tend to have a narrow focus on minimizing costs -- potentially at the expense of distributional concerns.
-The energy transition will have wide-ranging effects --such as shifts in the spatial distribution of jobs, or changes in air pollution, that are of interest to policy-makers, but are currently unincorporated into least-cost models.
+This paper takes a different approach.
+Instead of modifying the objective function to internalize monetized pollution damages, or analyzing expected health outcomes ex-post, we propose a new framework that endogenizes pollution exposure in the planning model and places distributional equity at the center of the optimization.
+Drawing inspiration from Rawlsian principles of justice, we formulate a min-max optimization in which the social planner minimizes the maximum pollution exposure across population groups, subject to an overall cost constraint.
+To do this, we use a reduced-form emissions to concentration model, InMAP, to calculate how much each unit of electricity generation in each model region affects population-weighted pollution exposure.
+Then, we use a least-cost electricity planning model, Switch, to solve for electricity system designs that achieve low pollution levels.
 
-Previous attempts to consider the health impacts of air pollution have monetized air pollution harms, and inserted them into the cost minimization.
-While this is certainly an improvement over not considering these costs at all, this method ignores the fact that inequity in the past has driven current differences in air pollution exposure, and that certain groups may experience higher burdens from air pollution than others.
+By solving this model at varying cost levels, we produce a set of frontier solutions that answer questions such as: How much can we equitably reduce pollution exposure for $\$$100 billion?
+This provides planners and policymakers with a tractable way to understand the trade-offs between cost, pollution improvements, and equity in electricity system design.
+Our approach could also generalize to other dimensions of equity.
+For example, it could be adapted to limit job losses in vulnerable regions.
+Prior research links declines in coal employment to a host of community-wide harms—such as reductions in home values, increases in poverty, and worsening credit outcomes—even for residents not directly employed in the industry (\citet{Kraynak2023TheCountry}; \citet{Blonz2023TheFuels}).
 
-However, in this paper, rather than modifying the objective function to include health effects, or use model outputs to project health changes, we endogenize pollution exposure within the electricity model in a way that explicitly considers distributional effects.
-We pull inspiration from Rawls and formulate our model as a min-max problem where the social planner minimizes the maximum pollution exposure subject to a cost constraint, rather than minimizing costs.
-Running multiple iterations of the model at different cost constraints allows us to understand how much can we equitably improve air quality and at what costs.
-
-This allows us to use the model to produce results that don't unfairly burden certain groups with disproportionate pollution exposure.
-It also can be used to answer questions such as "How much can pollution exposure be equitably reduced for $\$100$ billion?"
-
-This method could also be used to help design electricity transition plans that are equitable in other dimensions.
-For example, one concern with the energy transition is that some regions will lose jobs.
-This method could help policymakers and planners design electricity systems that avoid causing disruptive employment declines in certain regions.
-Previous work has linked declines in employment from  coal extraction to negative impacts such as decreases in home values, increases in poverty, and increases in credit utilization and delinquency.
-Moreover, these studies also suggest that even residents who are not employed by the industry also suffer the negatives consequences associated with the industry's decline (\citet{Kraynak2023TheCountry}, \citet{Blonz2023TheFuels}).
-
-This paper makes an important contribution to the environmental justice literature by ...
+This paper contributes to the environmental justice and electricity planning literatures by proposing a new modeling framework that explicitly considers the distribution of air pollution exposure.
+We cast an old way of thinking about society’s preference for equity – Rawls’s min-max, to solve a modern problem – fixing environmental justice issues resulting from legacy inequities.
+We show that substantial improvements in both aggregate exposures and exposure equity are achievable at modest additional system costs, and we provide a flexible tool for policymakers who aim to design a just energy transition.
 
 \section{Data}
-\subsection{Emissions Data}
+Data is required for two tasks in our method. The first task is to construct emissions rates for sources in the model in tons per MWh. The second task is to construct input data for Switch modelling.
+
+\subsection{Constructing Emissions rates}
+Emissions rates are constructed by combining emissions data from the U.S.
+Environmental Protection Agency (EPA), and generation data from the U.S.
+Energy Information Administration (EIA).
+We use the EPA-EIA Power Sector Data Crosswalk to merge generators across the two sources \citep{HuettemanEPA-EIACrosswalk}.
+
 Emissions data comes from two sources: EPA's Clean Air Markets Program Data (CAMD) and EPA's National Emissions Inventory (NEI).
@@
 I get emissions of NH$_3$, VOC, and PM2.5 from the NEI.
+Air pollution data is used to calculate source emissions rates in tons per MWh.
+
+Data on the generation and location of existing plants comes from the EIA. Generation data comes from EIA Form 923, and plant characteristic data comes from EIA Form 860.
 
 \subsection{Power Sector Modelling Inputs}
-Stuff about PowerGenome?
+To construct a tractable Switch model, we generate input data with PowerGenome, a data management platform that was originally developed to generate input datasets for GenX, another electricity planning model.
+PowerGenome relies on numerous public datasources such as the EIA, the Public Utility Data Liberation Project (PUDL), and the National Renewable Energy Lab (NREL).
+We discuss the datasources and data processing involved with Switch model input construction in section 3.3
 
@@
 \subsection{Overview}
+Our modelling can be broken down into four tasks.
+The first task is to generate input data for Switch modeling.
+The second task is to calculate resource-specific emissions rates, then generate coefficients that describe how one unit of electricity from each resource affects population weighted mean PM2.5 exposure for different groups.
+The thir task is to run Switch in cost-minimization mode.
+Once we've solved for minimum system costs, the final task is to re-run Switch in pollution minimization mode with a cost constraint.
+
+\subsection{Constructing Planning Model Inputs}
+To run Switch, we needed a set of input data.
+This modeling relies on input data and scenario construction from the Model Intercomparison Project.
+The Model Intercomparison Project utilizes PowerGenome, a data management platform that generates standardized input datasets for electricity sector planning models, to construct input data for different scenarios and configurations.
+PowerGenome harmonizes a wide range of public data sources and converts them into consistent, model-ready formats.
+These inputs include technology costs and performance, fuel prices, load forecasts, renewable profiles, policy incentives, and resource constraints.
+
+PowerGenome integrates generator-level operational and retirement data from the U.S.
+Energy Information Administration (EIA 860 and 860M) and the Public Utility Data Liberation (PUDL) project.
+Capital and operational costs are primarily drawn from the National Renewable Energy Laboratory's Annual Technology Baseline (NREL ATB).
+Load shapes are constructed using historical profiles from NREL, adjusted using sector-specific electrification projections from Princeton’s REPEAT project.
+Hourly wind and solar capacity factors are derived from Vibrant Clean Energy datasets.
+Fuel prices reflect the U.S.
+Energy Information Administration's Annual Energy Outlook.
+
+For the MIP project, inputs were configured to have a 26-zone spatial resolution over the continental United States, and operational choices were made using 10 representative weeks.
+In addition, to reduce the computational complexity of solving the model, resources are clustered by Powergenome within their model region and technology by heat rate and fixed operating costs.
+For example, ``Western ERCOT combined-cycle natural gas." For the remainder of the paper, we refer to these resource clusters as ``resources."
+
+We use the \textit{current policies} scenario, which incorporates all federal and state-level policies on the books as of 2023, including investment and production tax credits for eligible technologies, subsidies for carbon capture and hydrogen, and regional clean energy mandates.
+This scenario is policy-static, meaning no additional climate policy is assumed beyond what is already enacted.
+To reflect temporary policy instruments such as the Inflation Reduction Act (IRA) tax credits, PowerGenome amortizes the benefits of tax incentives over plant lifetimes and embeds them in the cost structure as reduced capital costs (for ITCs) or negative variable costs (for PTCs).
+Regional capacity mandates—such as offshore wind targets—and CO$_2$ caps are explicitly modeled where applicable.
+More details on PowerGenome and the current policies scenario can be found in the MIP paper \citep{Schivley2024ProcessModels}.
+
+\subsection{Air Pollution Modeling}
+The end product necessary to endogenize air pollution within electricity planning models is a set of coefficients that describe the resource-specific marginal effect of a unit of electricity generation on population-weighted mean pollution exposure for different racial groups (though you could also apply the same logic and methodology to differing socio-economic groups more broadly, or more fine slices of the population to more closely approximate a true min-max).
+
+To construct these coefficients, we begin by computing resource-specific emissions rates (based on empirical emissions and generation data).
+For many plants in the EIA datasets, we are able to directly compute plant-specific emissions rates by observing plant-specific annual generation and plant-specific emissions of each of the relevant pollutants.
+For plants that are missing entries, we assign them emission rates equal to the technology-specific generation-weighted mean emissions rates.
+Since the model is forward-looking we also needed to make modeling assumptions about what the emissions rates will be for plants that have yet to be constructed.
+For these resources, we simply assign them their technology-specific generation weighted means.
+The inputs data allows for natural gas plants with carbon capture and storage.
+Since none of these currently exist in the U.S., we make the assumption that their emissions rates are slightly higher than typical natural gas plants since the pollution control equipment increases those plant's heat rates (increases the amount of fuel needed to generate one unit of electricity).
+In the current policies scenario, no CCS plants are built anyways.
+However, if it turns out that pollution control equipment also removes some air pollutants, then our results would weakly understate how much you could reduce pollution for some budget, since CCS could look like a low-pollution, easily dispatchable source of electricity.
+
+Next, we downscale a model resource unit (e.g.
+1 MWh of generation from Western ERCOT combined cycle natural gas) to pounds of pollutants in specific locations.
+For existing resources, we do this by allocating a unit of generation to existing plants within region weighted for capacity.
+So, if there are two coal plants in model region A with equal capacities.
+Then one unit of electricity generation from coal in model region A will be equally divided between the two plants, then multiplied by each plant's emissions factors.
+For resources that have yet to be constructed (and thus, have no existing locations) we assume that their spatial distribution will be the same as existing resources.
+So, for new natural gas plants, we assign emissions to the same locations as existing and retired natural gas plants.
+This reflects an assumption that future siting patterns will be similar to past siting, and provides a conservative estimate of exposure.
+This highlights an important limitation of our model -- since model regions are relatively large, there is probably a lot of substitution \textit{within} model regions that could be done to reduce pollution that our results will not capture.
+However, this model limitation should bias the costs of reducing pollution \textit{upwards}, so our estimates of the costs of reducing pollution equitably are conservatively high.
+
+Finally, we use an emissions-to-exposure model to calculate how generation from each model resource affects group-level pollution exposure.
+Eulerian chemical transportation models are computationally expensive and time consuming, so it is infeasible to embed this type of model within Switch.
+Rather we use the InMAP Source-Receptor Matrix (ISRM) which is a source-receptor matrix approximation of the reduced form air transport model, InMAP, that describes the marginal effect of a unit of pollution on downwind concentrations of pollution exposure.
+
+InMap itself is a reduced-form emissions-to-exposure model that uses annual-average parameters to reduce computation complexity relative to a time-resolved chemical transport model.
+InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs the annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5 \citep{Tessum2017InMAP:Interventions}.
+The ISRM was developed to simplify InMAP even further \citep{Goodkind2019Fine-scaleEmissions}.
+Initially, the ISRM was developed to describe the marginal damages from individual sources.
+
+Recall that from our downscaling procedure, we have location specific emissions from 1 MWh of generation from each model resource.
+We then use the ISRM to calculate how each location affects population-weighted mean group exposure, then sum across all locations within a resource to generate a set of parameters describing how one unit of generation affects population-weighted group exposures measured in $\mu g/$MWh.\footnote{While our model focuses on equalizing population-weighted PM2.5 exposures across groups, it’s important to note that different populations may experience different health responses to the same exposure level.
+For example, due to preexisting health conditions, access to healthcare, housing quality, or cumulative environmental burdens, a given increase in PM2.5 may lead to more severe health outcomes in one group than another.
+\citet{Spiller2021MortalityOutcomes} show that using group-specific concentration response functions have significant effects on relative predicted mortality levels.} With those coefficients in hand, the Switch model, for each iteration of the optimization routine, multiplies those coefficients by resource-specific annual generation.
+
+\subsection{Switch Model}
+
+To model electricity system expansion under current policy constraints and distributional objectives, we use the Switch modeling platform.
+Switch is a modular, open-source power system optimization model designed to support long-term capacity expansion and detailed policy and economic analysis
+\citep{Johnston2019SwitchSystems}.
+Switch solves a linear or mixed-integer optimization problem that minimizes the net present value of investment and operational costs while meeting electricity demand and a variety of technical and policy constraints.
+The model operates on a structure of multiple investment periods (e.g., 2025, 2030, etc.), each of which contains representative days composed of hourly timepoints.
+This temporal structure enables Switch to model inter-hour dependencies essential for technologies like batteries, flexible loads, and thermal unit commitment.
+Switch supports detailed spatial modeling using multiple load zones connected by transport-constrained transmission corridors.
+
+Because Switch is implemented in Python using the Pyomo optimization modeling language, it is compatible with most commercial and open-source solvers.
+For this project we use Gurobi 12.0.1.
+Switch has been applied in diverse contexts including national-scale planning in Chile and China, subnational studies in Hawaii and the Western U.S., and evaluations of real-time pricing and distributed storage.
+For a detailed description of the software architecture and capabilities, see \citet{Johnston2019SwitchSystems}.
 
@@
         \begin{tikzpicture}[node distance=2cm]
-        \node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
-        • Constraints}};
+        \node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics}};
 
         \node(obj)[process, below of=start, text width=8cm]{\shortstack[l]{$\min \sum \text{Costs}$ \\ \textit{ 
-           s.t.   }\text{Everyone gets electricity and nothing breaks }}};
+           s.t.   }\text{Time-specific load demand is met and \\ physical generation and transmission constraints \\ are not violated}}};
 
@@
 
-\node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
-• Constraints}};
+\node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics}};
 
@@
 
-\subsection{Air Pollution Modelling}
-The end product necessary to endogenize air pollution within electricity planning models are a set of coefficients that describe the marginal effect of a unit of electricity generation on population-weighted mean pollution exposure for different racial groups (though you could also apply the same logic and methodology to differing socio-economic groups more broadly, or more fine slices of the population to more closely approximate a true min-max).
+\section{Results}
+\subsection{Model Inputs}
+There are a few findings from the construction of model inputs that are worth discussing.
+First, unsurprisingly, empirical data on emissions and generation reveals that coal is a dirty source of generation.
+However, when we translate emissions into marginal exposure coefficients, we find that the average relative difference between coal and natural gas decreases.
+In particular, it appears that natural gas plants are on average, closer to Asian and Latino population centers.
+A unit of electricity generated from coal only causes slightly more damage (in terms of population-weighted PM2.5 exposure) to Asians and Latinos than a unit of electricity generated from natural gas.
+In figure \ref{EmissionsExposureCoefs} we plot the average empirical emissions rates by technology in panel (a), and the mean coefficients describing how one MWh of generation affects mean group-level exposure of PM2.5 in $\mu$g/m$^3$.
 
-Eulerian chemical transportation models are computationally expensive and time consuming, so it's infeasible to embed this type of model within SWITCH.
-Rather we use the InMAP Source-Receptor Matrix (ISRM) which is a reduced-form source-receptor matrix that describes the marginal effect of a unit of pollution on downwind concentrations of pollution exposure.
-
-InMap itself is a reduced-form emissions-to-exposure model that uses annual-average parameters to reduce computation complexity relative to a time-resolved chemical transport model.
-InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs the annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5 \citep{Tessum2017InMAP:Interventions}.
-
-The ISRM was developed to simplify InMAP even further \citep{Goodkind2019Fine-scaleEmissions}.
-Initially, the ISRM was developed to describe the marginal damages from individual sources.
-We simplify the ISRM even further by collapsing it so that, for each source, we save the marginal effect of one MWh of generation on \textit{population-weighted average exposure} by race.
-
-\subsection{Switch}
-
-\section{Results}
+This stark divergence between the relative emissions rates and the relative exposure coefficients highlights one of the central premises of the paper -- that where generation is placed can have large impacts on pollution exposure, and that selectively choosing what generation goes where can lead to large decreases in pollution exposure at low costs.
 
@@
 \centering
-\subcaptionbox{Mean empirical emissions rates by technology in tons per MWh.}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/base_short_emissions_by_technology.png}}%
+\subcaptionbox{Mean empirical emissions rates by technology in tons per MWh.}{\includegraphics[width=0.8\textwidth]{Figures/EndogenousPaper/emissions_by_technology.png}}%
 \hfill
-\subcaptionbox{Mean coefficients describing how a marginal MWh of generation affects mean group-level PM2.5 exposure.}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/base_short_exposure_by_technology.png}}%
+\subcaptionbox{Mean coefficients describing how a marginal MWh of generation affects mean group-level PM2.5 exposure.}{\includegraphics[width=0.8\textwidth]{Figures/EndogenousPaper/exposure_by_technology.png}}%
 \hfill
-\caption{Unsurprisingly, coal is far dirtier than other technologies.  However, the discrepancy between technologies is more narrow for exposure.}
+\caption{Coal is far dirtier than other technologies when looking at emissions rates (Tons/MWh).
+However, the discrepancy between technologies is more narrow for exposure, reflecting differences across the technologies in their spatial distribution.}
 \label{EmissionsExposureCoefs}
 \end{figure}
-In figure \ref{EmissionsExposureCoefs} we plot both average emissions rates, and average exposure coefficients for various technologies.
+
+\subsection{Model Results}
+Our high-level, key finding is that even a modest budget of $\$50$ billion through 2050 could pay for massive reductions in equitable air quality improvement.
+For context, the total system costs in the base scenario are about $\$2.4$ trillion, so $\$50$ billion represents an increase in total costs of about 2$\%$.
+In the $\$50$ billion budget scenario, the maximum group level exposures are approximately cut in half in each model period relative to the base scenario.
+
+Our results have features that you would expect based on economic theory.
+In figure \ref{PPF}, we plot the maximum group exposure in each model period against the budget for reducing pollution.
+There are increasing costs of reducing air pollution -- in other words, the gains from moving from least-cost to allocating $\$50$ billion to reduce pollution exposure are much larger than the gains from moving from allocating $\$50$ billion to $\$100$ billion.
 
@@
     \centering
-    \subcaptionbox{2027}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/Dispatch_Relative_by_scenario_2027.png}}
-    \subcaptionbox{2030}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/Dispatch_Relative_by_scenario_2030.png}}
-    \caption{Generation by Technology by Scenario}
+    \includegraphics[width=0.7\linewidth]{Figures/EndogenousPaper/exposure_cost_PPF.png}
+    \caption{The relationship between the objective function (min-max of exposure in $\mug$/MWh) and the cost limit constraint.}
+    \label{PPF}
 \end{figure}
+
+In figure \ref{ExposureByRace} we've plotted mean exposure by race across different budgets in the year 2050 model period.
+Our results show that the groups with the highest baseline exposures -- Black, Asian, and Latino Americans experience the greatest gains from the scenarios which reduce air pollution.
+Our results also suggest that at the optimum, the min-max objective generates equal average exposures for most groups.
+In other words, the min-max function binds for most groups.
 
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{Figures/EndogenousResults/base_short/exposure_cost_PPF.png}
-    \caption{Here we show the relationship between the objective function (min-max of exposure) and the cost limit constraint.}
-    \label{PPF}
+    \includegraphics[width=0.5\linewidth]{Figures/EndogenousPaper/exposure_by_scenario_2050.png}
+    \caption{Population-weighted mean air pollution exposure by race in 2050 across budget scenarios.}
+    \label{ExposureByRace}
 \end{figure}
 
+\subsection{Mechanisms}
+
+The model primarily produces air pollution benefits by shifting away from fossil generation towards zero-emissions, renewable generation.
+In figure \ref{Generation}, we've plotted annual generation from different technologies, by budget scenario and model year.
+
+\begin{figure}
+    \centering
+    {\includegraphics[width=0.45\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2027.png}}
+    {\includegraphics[width=0.45\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2030.png}}\\
+    {\includegraphics[width=0.45\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2035.png}}
+    {\includegraphics[width=0.45\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2040.png}}\\
+    {\includegraphics[width=0.45\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2045.png}}
+    {\includegraphics[width=0.45\textwidth]{Figures/EndogenousPaper/Dispatch_Relative_by_scenario_2050.png}}\\
+    \includegraphics[width=0.45\textwidth]{Figures/EndogenousPaper/Legend_2050.png}
+    \caption{Generation from each technology category across scenario-years.}
+    \label{Generation}
+
+\end{figure}
+
+In the nearer-term, the model chooses to replace almost all existing coal generation with natural gas and wind generation.
+Note that in 2050, even in the base scenario, the cost-minimizing model almost entirely pushes out coal.
+So, one major thing the model spends its budget on when trying to reduce pollution exposure is getting rid of coal sooner, especially in the mid-Atlantic and Midwest regions.
+We've plotted the difference between generation from coal in the base and $\$200$ billion scenario in figure \ref{coalmap}.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{Figures/EndogenousPaper/coal_generation_map_2027_200B.png}
+    \caption{Difference in coal generation in 2027 with $\$200$ billion budget relative to the base scenario}
+    \label{coalmap}
+\end{figure}
+
+In the far-term, the model broadly chooses to replace natural gas with wind and solar.
+In figure \ref{2050Generation} we plot differences in natural gas, wind, and solar generation.
+However, rather than decreasing natural gas use across the board, the model instead chooses to sharply reduce natural gas generation in some regions, such as eastern Texas, southern California, the Midwest, and the northern Mid-Atlantic, while modestly increasing natural gas generation in other areas like west Texas and the Pacific Northwest.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{Figures/EndogenousPaper/naturalgas_generation_map.png}
+    \includegraphics[width=0.8\linewidth]{Figures/EndogenousPaper/wind_generation_map.png}
+    \includegraphics[width=0.8\linewidth]{Figures/EndogenousPaper/solar_generation_map.png}
+    \caption{Differences in generation from natural gas, wind, and solar in 2050 with a $\$200$ billion budget relative to the base scenario}
+    \label{2050Generation}
+\end{figure}
+
+In a few areas, the model uses transmission as a way to decrease air pollution.
+In figure \ref{transmissionsubstitution}, we've plotted an example of how this works.
+To reduce air pollution, the model over-constructs capacity in some regions, in this example, southern part of the Southwest Power Pool (SPP), and allows that region to export to adjacent areas.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{Figures/EndogenousPaper/transmission_capacity_substitution.png}
+    \caption{Differences in Transmission and Generation Capacity Relative to the Base Scenario}
+    \label{transmissionsubstitution}
+\end{figure}
+
+\subsection{Discussion}
+While we think these results have important policy implications, namely that pollution exposure can be dramatically reduced at low costs, it's important to note that our results don't point to which policies would achieve the optimal solution.
+Rather, these results are prescriptive -- \textit{"Build X, Y, and Z in regions A,B, and C to achieve the deepest equitable cuts in air pollution at a certain cost."} However, these results could point towards policies which might approximate model results, such as targeted plant retirements or regional clean energy standards.
+
+While our results are new, we think it's important to contextualize them with other findings from the literature.
+For example, \citet{Shawhan2024PoliciesAmericans} and \citet{Goforth2022AirStrategies} find that the U.S.
+is on track to dramatically reduce pollution exposure, and that faster decarbonization policies will generate deep cuts in air pollution exposure.
+Our results are broadly consistent with these findings.
+For example, in the base, current policies scenario, we predict that Black Americans will experience the largest drop in pollution exposures over the study period-- a decrease of about 50$\%$ between 2027 and 2050.
+
+These results are somewhat unsurprising given the tight link between carbon emissions and pollution emissions. In figure \
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{Figures/EndogenousPaper/emissions_vs_exposure.png}
+    \caption{Caption}
+    \label{emissionsexposure}
+\end{figure}
 \section{Conclusion}
-In this paper, we demonstrate how to endogenize air pollution exposure into a least-cost electricity planning model to create equitable reductions in air pollution burdens.
+This paper develops and applies a novel framework for integrating air pollution equity into electricity planning models.
+Rather than optimizing purely for cost or incorporating pollution as a monetized externality, we directly embed pollution exposure—measured in terms of group-specific PM2.5 burden—into the objective function using a Rawlsian min-max approach.
+We combine empirical emissions data, an atmospheric dispersion model (InMAP), and a least-cost power system optimization model (Switch) to identify power sector investments that minimize the maximum pollution burden across racial groups, all while remaining within a specified system cost constraint.
 
-While this method provides a map of how to produce equitable reductions in air pollution, it does not solve for what policies will get
+Our results demonstrate that substantial improvements in both aggregate and distributional air quality outcomes are achievable with relatively modest expenditures.
+For example, a $\$50$ billion investment—spread across several decades—can cut maximum group-level exposures in half relative to the least-cost baseline.
+The gains are not evenly distributed: Black, Latino, and Asian populations, who face the highest baseline exposures, benefit most from the model’s pollution-reduction strategies.
+Mechanistically, these improvements are achieved by accelerating coal retirements, strategically reducing natural gas generation in densely populated or upwind areas, increasing renewable deployment, and upgrading transmission infrastructure to allow for cleaner electricity imports.
+
+Importantly, our modeling framework is prescriptive, not predictive.
+It does not identify specific policies, incentives, or mandates that would deliver the modeled outcomes.
+Instead, it provides a decision-support tool to help planners and policymakers understand the cost–equity tradeoffs embedded in electricity system design.
+As such, it complements—but does not replace regulatory or political processes.
+
+This framework is generalizable.
+While we focus on air pollution equity, the same structure could be used to minimize job losses in vulnerable communities, optimize energy affordability across income groups, or equitably distribute benefits like clean energy jobs or infrastructure investment.
+Future work could extend this framework to include multiple dimensions of equity simultaneously or explore how it interacts with climate targets and carbon pricing.
+In sum, our work shows that by modifying standard electricity planning tools to account for pollution exposure and equity, it is possible to design cleaner, more equitable power systems at a reasonable cost.
 
@@
 
+\section{Appendix}
 \end{document}
